### 2022-11-15 test.Rmd
edited test.qmd
added new joined csv

---
title: "test"
output: html_document
date: "2022-11-06"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
library(lubridate)
```

## Read in data

```{r}
# Read in soundscape data, round to nearest 20 min
BB_08_csv <- read_csv("Data/CCES_08_1Hz_1s_BB_2min.csv")
BB_08 <- BB_08_csv %>% 
  rename(dateTime = `yyyy-mm-ddTHH:MM:SSZ`)
BB_08 <- BB_08 %>%
  mutate(dateTime = round_date(ymd_hms(BB_08$dateTime),"20 minutes"))
```

### Read in Tracks
```{r}
# Read in tracks, filter out lost buoys
tracks_csv <- read_csv("Data/AllTracks.csv")
tracks <- tracks_csv %>%
  dplyr::filter(station == "4"| station == "7"| station =="8" |station == "10"| 
                station == "12"|station == "13"| station == "14"| station == "16"|
                station == "17"|station == "18"| station == "19"| station == "20"| 
                station == "21"|station == "22"| station == "23")

# Filter by CCES_08 buoy
tracks_08 <- tracks %>% dplyr::filter(station == "8")

# Round CCES_08 to nearest 20 min
tracks_08 <- tracks_08 %>% 
  mutate(dateTime = ceiling_date(ymd_hms(tracks_08$dateTime), "20 minutes"))

```

### Read in OL
```{r}
# Read in OL data, round to nearest 20 min
OL_08_csv <- read_csv("Data/CCES_08_1Hz_1s_OL_2min.csv")
OL_08 <- OL_08_csv %>% 
  rename(dateTime = `yyyy-mm-ddTHH:MM:SSZ`)
OL_08 <- OL_08 %>%
  mutate(dateTime = round_date(ymd_hms(OL_08$dateTime),"20 minutes"))
OL_08

```

Joining tables
```{r}
# Join tables
track_BB_08 <- left_join(tracks_08_rnd, BB_08, by = "dateTime")
track_BB_OL_08 <- left_join(track_BB_08, OL_08, by="dateTime")

write_csv(track_BB_OL_08, "Data/track_bb_OL_08.csv")
```