### 2023-02-07 manuscript.qmd

---
title: "manuscript"
format: html
editor: visual
---

# Manuscript Draft

## Introduction

## Methods

Acoustic detections were collected with the use of free-drifitng Drifting Acoustic Spar Buoy Recorders (DASBRs). These were deployed as part of the California Current Ecosystem Survey (CCES) spanning the offshore waters from southern Canada to northern Mexico in summer of 2018. DASBRs offer greater spatial and temporal coverage compared to bottom-mounted or stationary hydrophones, ideal for detection of offshore or mobile species. These DASBRs contained two satellite geo-location trackers, two hydrophones, etc and collected a total of 1910 hours of data. SWFSC ran acoustic detections in PAMGuard. Eight drifting recorders were lost at sea and therefore removed from the analysis. These contain 1, 5, 10, 25, 75, 95, and 99 percentile distribution from each metric

```{r}

```

## Results

## Discussion

```{r}
#| echo: false

```

The `echo: false` option disables the printing of code (only output is displayed).