### 2023-02-07 test.Rmd

---
title: "test"
output: html_document
date: "2022-11-06"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
library(lubridate)
library(sf)
library(rerddap)
library(terra)
```

## Read in CCES08 data

```{r}
# Read in soundscape data, round to nearest 20 min
BB_08_csv <- read_csv("Data/CCES_08_1Hz_1s_BB_2min.csv", show_col_types = FALSE)
BB_08 <- BB_08_csv %>% 
  rename(dateTime = `yyyy-mm-ddTHH:MM:SSZ`)
BB_08 <- BB_08 %>%
  mutate(dateTime = round_date(ymd_hms(BB_08$dateTime),"20 minutes"))

```

### Read in Tracks

```{r}
# Read in tracks, filter out lost buoys
tracks_csv <- read_csv("Data/AllTracks.csv", show_col_types = FALSE)
tracks <- tracks_csv %>%
  dplyr::filter(station == "4"| station == "7"| station =="8" |station == "10"| 
                station == "12"|station == "13"| station == "14"| station == "16"|
                station == "17"|station == "18"| station == "19"| station == "20"| 
                station == "21"|station == "22"| station == "23")

# Filter by CCES_08 buoy
tracks_08 <- tracks %>% dplyr::filter(station == "8")

# Round CCES_08 to nearest 20 min
tracks_08_rnd <- tracks_08 %>% 
  mutate(dateTime = round_date(ymd_hms(tracks_08$dateTime), "20 minutes"))
tracks_08_rnd
```

### Read in OL, round to nearest 20 min

```{r}
# CCES_08
OL_08_csv <- read_csv("Data/CCES_08_1Hz_1s_OL_2min.csv")
OL_08 <- OL_08_csv %>% 
  rename(dateTime = `yyyy-mm-ddTHH:MM:SSZ`)
OL_08 <- OL_08 %>%
  mutate(dateTime = round_date(ymd_hms(OL_08$dateTime),"20 minutes"))
OL_08

# CCES_10
OL_10_csv <- read_csv("Data/CCES_10_1Hz_1s_OL_2min.csv")
OL_10 <- OL_10_csv %>% 
  rename(dateTime = `yyyy-mm-ddTHH:MM:SSZ`)
OL_10 <- OL_10 %>%
  mutate(dateTime = round_date(ymd_hms(OL_10$dateTime),"20 minutes"))
OL_10
```

### Read in CCES_10

```{r}
# Read in CCES_10 soundscape metrics
BB_10_csv <- read_csv("Data/CCES_10_1Hz_1s_BB_2min.csv")
BB_10 <- BB_10_csv %>% 
  rename(dateTime = `yyyy-mm-ddTHH:MM:SSZ`)
BB_10 <- BB_10 %>%
  mutate(dateTime = round_date(ymd_hms(BB_10$dateTime),"20 minutes"))

# Filter by CCES_10 buoy
tracks_10 <- tracks %>% dplyr::filter(station == "10")

# Round CCES_10 to nearest 20 min
tracks_10_rnd <- tracks_10 %>% 
  mutate(dateTime = ceiling_date(ymd_hms(tracks_10$dateTime), "20 minutes"))

```

### Read in whale detections

```{r}
# Read in whale detections from CCES_08 and CCES_10
whales <- read_csv("Data/EventInfo.csv")
whales_sf <- whales %>%
  dplyr::filter(Deployment == "8"| Deployment == "10") %>%
  st_as_sf(coords = c("Longitude","Latitude"), crs=4326) %>%
  mutate(bestNumber = as.numeric(bestNumber)) %>%
  mutate(nClicks = as.numeric(nClicks))

```

## Read in AIS Data
```{r}
rawAIS<- read_csv("Data/AIS_072518_072918.csv")
# need to figure out where to store data
```

### Joining tables with soundscape metrics

```{r}
# CCES_08
track_BB_08 <- left_join(tracks_08_rnd, BB_08, by = "dateTime")
track_BB_OL_08 <- left_join(track_BB_08, OL_08, by="dateTime")
write_csv(track_BB_OL_08, "Data/track_bb_OL_08.csv")

# CCES_10
track_BB_10 <- left_join(tracks_10_rnd, BB_10, by = "dateTime")
track_BB_OL_10 <- left_join(track_BB_10, OL_10, by="dateTime")
write_csv(track_BB_OL_10, "Data/track_bb_OL_10.csv")
```

### Map Animations
```{r}
# Install
#devtools::install_github("ropenscilabs/rnaturalearth")
#install.packages("rnaturalearth",
                 #repos = "http://packages.ropensci.org",
                 #type = "source")
#install.packages("gganimate")
#install.packages("gifski")
#install.packages("marmap")

# Libraries
library(sf)
library(tidyverse)
library(marmap)
library(gifski)
library(gganimate)
library(rnaturalearth)
```

### New ggplot wtih bplot as basemap 
```{r}
library(mapproj)
# Get bathymetry (code from Ocean Tracking Network https://www.youtube.com/watch?v=xT2F-Y9bKLk)
bgo <- getNOAA.bathy(lon1 = min(df_all$lon-0.7), lon2 = max(df_all$lon + 0.7),
                     lat1 = min(df_all$lat-0.7), lat2=max(df_all$lat + 0.7), resolution = 1)

# Create basemap
bf <- fortify.bathy(bgo, geom=c("r", "c")) + scale_fill_etopo() # converts bathy object to data frame
bplot <- autoplot(bgo, geom = c("raster", "contour"), coast = TRUE) + 
  scale_fill_etopo() #creates ggplot object

# Plot
p <- bplot + 
  geom_path(data = track_BB_OL_10, 
            aes(x=long,y=lat,group=spotID), 
            alpha = 0.5,
            linewidth = 3) + # line width parameter is ignored for some reason
  geom_point(data = track_BB_OL_10, 
             aes(x=long,y=lat,group=spotID, size = OL_125, color = OL_125),
             shape=19, stroke = 4) +
  #formatting  
  scale_color_viridis_c(option = "inferno") +
  scale_size_continuous(range = c(0.1,10)) +
  labs(x=NULL, y=NULL, 
       fill = 'Depth',
       color = 'OL 125')+
  theme(panel.grid = element_blank())
```

```{r}
# Animate
anim = p +
  transition_reveal(along = dateTime)+
  ease_aes('linear')+
  shadow_wake(0.1) +
  ggtitle("Date: {frame_along}")

# Animate with control over frames per second and number of frames
animate(anim, nframes = 70, fps = 3)

```

Linear Regression Model

```{r}
# Does sound vary with latitude?
lm(formula = OL_63 ~ lat, data = track_BB_OL_08)
fit10 <- lm(formula = lat ~ OL_63, data = track_BB_OL_10)
plot(fit10)
ggplot(track_BB_OL_08, aes(x = lat, y = `BB_20-24000`)) + scale_y_continuous() +
   geom_point() + geom_smooth()

# Does sound vary with date/time?
summary(lm(formula = `BB_20-24000`~ dateTime, data = track_BB_OL_10))

ggplot(track_BB_OL_08, aes(x = dateTime, y = `BB_20-24000`)) +
  geom_point() +
  geom_smooth() +
  ggtitle("Buoy 08 - Sound levels by date")

ggplot(track_BB_OL_10, aes(x = dateTime, y = `BB_20-24000`)) +
  geom_point() +
  geom_smooth() +
  ggtitle("Buoy 10 - Sound levels by date")

plot(lm(formula = `BB_20-24000`~ dateTime, data = track_BB_OL_08))

ggplot() +
  geom_point(data = track_BB_OL_08, aes(x = dateTime, y = `BB_20-24000`, color = "red")) +
  geom_point(data = track_BB_OL_10, aes(x = dateTime, y = `BB_20-24000`, color = "blue"))

```

CCES Broadband and OL Map

```{r}
library(tmap); library(maptiles); library(sf)
tmap_mode("plot")
trackssf <- st_as_sf(tracks, coords = c("long", "lat"), crs=4326)
tracks08sf <- st_as_sf(track_BB_OL_08, coords = c("long", "lat"), crs=4326)
tracks10sf <- st_as_sf(track_BB_OL_10, coords = c("long", "lat"), crs=4326)
trackBase <- get_tiles(whales_sf, provider = "Esri.OceanBasemap")
allTrackbase <- get_tiles(trackssf, provider = "Esri.OceanBasemap")

#Map all tracks
tm_shape(allTrackbase) + tm_rgb() +
  tm_shape(trackssf) + 
  tm_dots(col = "black")

# Map tracks 08 and 10 only
tm_shape(trackBase) + tm_rgb() + 
  tm_shape(tracks08sf) + tm_dots(col = "red", 
                                 legend.show = TRUE) +
  tm_shape(tracks10sf) + tm_dots(col = "green")

# CCES_08 + CCES_10 Broadband
tm_shape(trackBase) + tm_rgb() + 
  tm_shape(tracks08sf) + 
  tm_symbols(col = "BB_20-24000", 
             border.lwd = NA, 
             size = 0.4, 
             sizes.legend = 0.7, 
             n = 10) +
  tm_shape(tracks10sf) + 
  tm_symbols(col = "BB_20-24000",
             border.lwd = NA, 
             size = 0.4,
             n = 10) +
  tm_layout(legend.width = 0.5, 
            legend.position = c("right","top"),
            legend.bg.color = "lightblue")

#CCES_08 Octave Level
tm_shape(trackBase) + tm_rgb() + 
  tm_shape(tracks08sf) + 
  tm_symbols(col = "OL_63", 
             border.lwd = NA, 
             size = 0.5, 
             n = 10) + 
  tm_layout(legend.width = 0.5, 
            legend.position = c("right","top")) + 
  tm_legend()

```

### Sea Surface Temp with rerddap

```{r}
# Looking at one day, Aug 23 2018

library(rerddap)
SSTInfo <- info('jplMURSST41')
murSST <- griddap(sstInfo, latitude = c(22., 51.), longitude = c(-140., -105), time = c("2018-08-23", "2018-08-23"), fields = 'analysed_sst')
ggplot(data = murSST$data, aes(x = longitude, y = latitude, fill = analysed_sst))

# needs fixing!

# Trying SST here again...
library(terra)
SST <- terra::rast("Data/SST.tif")
track_rast <- rasterize(tracks08sf, SST)

plot(SST)
plot(track_rast)
```

### Bathymetry Analysis

```{r}
# Increase cell size by 30x to be visible
bathy <- rast("Data/gebco_bathy.tif")
agg <- terra::aggregate(bathy, 30)
plot(agg)

# Rasterize track with agg as the template
track_rast <- rasterize(tracks08sf, agg, field = "BB_20-24000")
plot(track_rast)

bathy_df <- as.data.frame(bathy, xy = TRUE)

# Map the coastline
cont <- st_as_sf(as.contour(bathy))
coast <- cont %>% filter(level == 0)
cont_slope <- cont %>% filter(level == -2000)
cont_slope_rast <- rasterize(cont_slope, agg, field = "level")
plot(cont_slope_rast, col = "red")
cont_slope_rast <- rasterize(cont_slope, agg)
plot(cont_slope_rast, col = "red")

ggplot() + 
  geom_raster(data = bathy_df, 
              mapping = aes(x=x, y=y, 
                            alpha = 0.8,
                            fill = `gebco_bathy`))+
  geom_sf(data = cont, col="black") + 
  geom_sf(data = coast, col = "red") +
  geom_sf(data = cont_slope, col = "purple") +
  geom_sf(data = tracks08sf, mapping = aes(col= `BB_20-24000`)) +
  geom_sf(data = tracks10sf, mapping = aes(col= `BB_20-24000`)) +
  scale_color_gradient(low="green", high="red")

# Distance to continental slope

BB_dist <- terra::distance(track_rast, vect(cont_slope))
dist_ex08 <- terra::extract(BB_dist, vect(tracks08sf))
pts08 <- cbind(tracks08sf,dist_ex08) %>%
  dplyr::select(-ID) %>%
  rename("Distance to Continental Slope" = gebco_bathy)

dist_ex10 <- terra::extract(BB_dist, vect(tracks10sf))
pts10 <- cbind(tracks10sf,dist_ex10) %>%
  dplyr::select(-ID) %>%
  rename("Distance to Continental Slope" = gebco_bathy)

# Map distance to continental slope by color
ggplot() +
  geom_sf(data = pts08, 
          mapping = aes(col = `Distance to Continental Slope`)) +
  geom_sf(data = pts10, 
          mapping = aes(col = `Distance to Continental Slope`)) +
  geom_sf(data = cont_slope)

# Plot distance by BB 
ggplot(data = pts08, aes(x = `Distance to Continental Slope`, y = `BB_20.24000`)) + geom_boxplot()

summary(lm(formula = BB_20.24000 ~ `Distance to Continental Slope`, data = pts08))

```

Does hour of day affect sound levels?

```{r}
# CCES_08
hr_08 <- track_BB_OL_08 %>%
  mutate(Hr = hour(track_BB_OL_08$dateTime)) %>%
  group_by(Hr) %>%
  summarize(meanBB = mean(`BB_20-24000`)) %>%
  rename(`Buoy 08 Broadband` = `meanBB`)
ggplot(data = hr_08, mapping = aes(x = Hr, y = `Buoy 08 Broadband`)) + 
  geom_point() + geom_smooth()

#CCES_10
hr_10 <- track_BB_OL_10 %>%
  mutate(Hr = hour(track_BB_OL_10$dateTime)) %>%
  group_by(Hr) %>%
  summarize(meanBB = mean(`BB_20-24000`)) %>%
  rename(`Buoy 10 Broadband` = `meanBB`)
ggplot(data = hr_10, mapping = aes(x = Hr, y = `Buoy 10 Broadband`)) + 
  geom_point() + geom_smooth()

# Join and graph
hr_join <- left_join(hr_08, hr_10, by = "Hr")
hr_piv <- hr_join %>%
  pivot_longer(col=c(`Buoy 08 Broadband`,`Buoy 10 Broadband`), names_to = "buoys", values_to = "BB")

ggplot(data = hr_piv, mapping = aes(Hr, BB, col=buoys)) +
  geom_point() + geom_smooth()
```

### Are sound levels affected by species composition or number of clicks?

```{r}
# Plot
library(tmap); library(maptiles)
tmap_mode("plot")
oceanBase <- get_tiles(whales_sf, provider="Esri.OceanBasemap", crop = TRUE)
tm_shape(oceanBase) + tm_rgb() +
  tm_shape(whales_sf) + 
  tm_symbols(col="species", size = "nClicks") +
  tm_layout(legend.width = 0.5, 
            legend.position = c("right","top"))

# CCES 08
BB_whales08 <- st_join(pts08, whales_sf) %>%
  filter(!is.na(`species`))

summary(glm(`BB_20.24000` ~ species + `Distance to Continental Slope` + nClicks, data = BB_whales08))

# CCES 10
BB_whales10 <- st_join(pts10, whales_sf) %>%
  filter(!is.na(`species`))

summary(glm(`BB_20.24000` ~ species + `Distance to Continental Slope` + nClicks, data = BB_whales10))

#Raster base map
tmap_mode("plot")
oceanBase <- get_tiles(BB_whales10, provider="Esri.OceanBasemap")
tm_shape(bathy) + tm_raster() +
  tm_shape(BB_whales10) + 
  tm_symbols(col="species", size = "nClicks", style = "jenks") +
  tm_layout(legend.width = 0.5)

```