### 2023-03-08 outline.qmd
updated with progress

---
title: "Thesis outline"
format: docx
editor: visual
---

## Goal 1: Characterize the soundscape

#### Explore the relationship between soundscape metrics and environmental variables, and beaked whale (BW) detections

-   Create histograms, scatterplots, time series, facet graphs of:
    -   A soundscape metric/BW presence + absence/env variable
        -   Comparing TOL intensities between drift 8 and 10
        -   Broadband levels, Third-octave levels at 63, 125, 2000, 5000, 20000 Hz
        -   Probability of detecting BW is highest when soundscape is \_\_\_.
        -   Which frequencies had the most loudness?
    -   Enviromental variables (source from PAMpal for events or ERDAPP)
        -   Bottom depth (**depth for whale detections in drift 8**)

        -   Bottom slope (**done for drift 8**)

        -   Distance from shore **(done for drift 8)**

        -   Distance from continental shelf (**done for drift 8**)

        -   Sea surface temperature (**in progress)**

        -   Sea surface height (need help sourcing, can only find SSH anomalies for 2018)

        -   Sea surface slope

        -   Chlorophyll-a (**mapped for drift 8)**

        -   Mixed layer depth (MLD), thermocline temperature (TTemp), temperature at 400 m, salinity (**Taiki's code)**
    -   Other variables: seasonality, hour of the day (create time series for avg daily env.variable)
    -   Questions: Does soundscape change when conditions change? Is sound seasonal?
-   Plug into model (GAM, zero-inflated, GLM)

## Goal 2: Influence from vessels

#### Explore how vessel presence influences the soundscape

-   Find instances of vessel in vicinity of drift vs. drift alone and compare

    -   Number of ship passages

        -   Create buffer around whale detections (unsure of buffer size)
        -   Find AIS vessels by lat/long that lie within buffer
        -   Quantify number of vessels that pass for each drift

    -   Closest ship passage (When a ship has a close encounter with a drift, how close is it on average? How is soundscape affected?)

-   Is vessel presence = loud soundscape?

    -   Compare TOL125 intensities vs. ship presence

-   Chronic vs. acute sound events

    -   Instances where drifts and vessels overlap, plot time on x-axis and TOL125 dB on y