### 2023-03-08 hycomThermocline[2998].rmd
Create hycomThermocline[2998].rmd

---
title: "HYCOM Data Access and Thermocline Variables"
author: "Taiki Sakai"
output: html_document
editor_options: 
  markdown: 
    wrap: 72
---

Prep some fake data. Environmental download functions require columns
named exatly `Longitude`, `Latitude`, and `UTC`. `UTC` must be converted
to POSIXct format.

```{r}
data <- data.frame(Latitude = c(38, 22.2, 15),
                   Longitude = c(-123.5, -152.9, 144),
                   UTC = c('2022-09-27 19:00:48',
                           '2022-09-27 17:57:48',
                           '2018-07-20 20:49:00'))
data$UTC <- as.POSIXct(data$UTC, format='%Y-%m-%d %H:%M:%S', tz='UTC')
```

Install latest PAMmisc from GitHub

```{r, eval=FALSE}
# only run this chunk once
devtools::install_github('TaikiSan21/PAMmisc')
```

## HYCOM Download

HYCOM has different datasets depending on the date, by passing this
`hycomList` object `PAMmisc` will sort out which one you need based on
your date ranges in your data, even figuring out if your download needs
to span multiple datasets.

The Hycom servers can be a bit slow, so this may take some time

```{r, cache=TRUE}
library(PAMmisc)
hycomData <- matchEnvData(data, nc=PAMmisc::hycomList, var=c('water_temp', 'salinity'), raw=TRUE, depth = 0:1000)
```

Normally the `matchEnvData` function will just add the environmental
variables to your dataframe, averaging them over any depth values. By
setting `raw=TRUE`, we instead have it return all values at the depths
so that we can work with them. The output is a list with length equal to
the number of rows in our dataframe.

```{r}
print(hycomData[[1]])
```

The matchXXX data show the coordinate values in the NetCDF files that
your data were matched to. The function will match your data to the
closest value in the NetCDF file.

## Thermocline Varibales

Now lets use the temperature at depth data to estimate thermocline
variables. This function returns estimates of mixed layer depth
(`mldDepth`), mixed layer temperature (`mldTemp`), thermocline depth
(`ttDepth`), and thermocline temperature (`ttTemp`). You don't need to
worry about the xxxMode parameters, they are place holders in case we
want to add different estimation methods in the future.

```{r}
calcThermoVars <- function(depth, temp, mldMode='SST8', thermMode='variso', plot=FALSE) {
    # depthInterp <- approxfun(x=temp, y=depth)
    switch(match.arg(mldMode),
           'SST8' = {
               mldTemp <- temp[1] - 0.8
               # first few temps are likely to have repeat values that trigger
               # warnings for approxfun, so cut off to only values after mldTemp
               useMin <- min(which(temp <= mldTemp)) - 1
               useIx <- useMin:length(temp)
               depthInterp <- approxfun(x=temp[useIx], y=depth[useIx])
               mld <- depthInterp(mldTemp)
           }
    )
    switch(match.arg(thermMode),
           'variso' = {
               t400 <- approx(x=depth, y=temp, xout=400)$y
               # therm end is half way to t400 temp
               tEnd <- mldTemp + (t400 - mldTemp)/2
               # therm temp is midpt from mld to end
               tt <- (mldTemp + tEnd)/2
               # tt <- mldTemp - .25 * (mldTemp - t400)
               ttDepth <- depthInterp(tt)
               tEndDepth <- depthInterp(tEnd)
           }
    )
    if(plot) {
        to300 <- depth <= 300
        plot(x=temp[to300], y=-depth[to300],
             main = paste0('MixLayer: ', round(mld,0), 'm ', round(mldTemp, 1), 'C',
                           '\nTherm: ', round(ttDepth, 0), 'm ', round(tt, 1), 'C'),
             xlab='Temp (C)', ylab='Depth (m)')

        lines(x=temp[to300], y=-depth[to300])
        lines(x=c(mldTemp, tEnd), y=c(-mld, -tEndDepth), col='darkgray', lwd=4)
        points(x=c(mldTemp, tt), y=c(-mld, -ttDepth), col='blue', cex=1.5, pch=15)
    }
    list(mldDepth = mld, mldTemp = mldTemp,
         ttDepth = ttDepth, ttTemp = tt)
}
```

We can calculate these and plot the results for our first data point.
The points are the temperature values at the depths in the HYCOM data.
The light gray line is the estimated thermocline based on the
calculations in [Fiedler
2010](https://aslopubs.onlinelibrary.wiley.com/doi/epdf/10.4319/lom.2010.8.313).
The blue squares mark the calculated mixed layer depth and thermocline
depth.

```{r}
calcThermoVars(hycomData[[1]]$matchDepth, hycomData[[1]]$water_temp, plot=TRUE)
```

And here's how we would connect these to our original dataframe, adding
in the temp & salinity at 400m. The list `hycomData` is in the same
order as the original rows, so we can just `cbind` them.

```{r}
library(dplyr)
thermData <- bind_rows(lapply(hycomData, function(x) {
    # do thermo calcs
    result <- calcThermoVars(x$matchDepth, x$water_temp)
    # get specific values
    result$temp400 <- x$water_temp[x$matchDepth == 400]
    result$sal400 <- x$salinity[x$matchDepth == 400]
    result
}))
data <- cbind(data, thermData)
print(data)
```