### 2023-06-29 07_summary_stats.Rmd
Create 07_summary_stats.Rmd

---
title: "07_summary_stats"
author: "Marina Bozinovic"
date: "2023-06-28"
output: html_document
---

```{r setup, include=FALSE}
library(tidyverse)
```

Read in RDS files of each buoy + all environ variables + AIS (pending)
```{r}
ftlist <- dir("data/", recursive=TRUE, full.names=TRUE, pattern="fulltrack")

for (ft in ftlist) {
  df <- readRDS(ft)
  assign(paste0("fulltrack", substr(ft, 15, 16)), df)
}
```

Read in RDA files for percentiles
```{r}
## Make list of rda files to read in
pclist <- c("BB_pct01_07", "BB_pct01_08", "BB_pct01_10", "BB_pct01_12", "BB_pct01_13", "BB_pct01_14", "BB_pct01_16", 
            "BB_pct01_18", "BB_pct01_19", "BB_pct01_20", "BB_pct01_21", "BB_pct01_22", "BB_pct01_23", "TO_pct01_07",
            "TO_pct01_08", "TO_pct01_10", "TO_pct01_12", "TO_pct01_13", "TO_pct01_14", "TO_pct01_16", "TO_pct01_18", 
            "TO_pct01_19", "TO_pct01_20", "TO_pct01_21", "TO_pct01_22", "TO_pct01_23", "BB_pct99_07", "BB_pct99_08", 
            "BB_pct99_10", "BB_pct99_12", "BB_pct99_13", "BB_pct99_14", "BB_pct99_16", "BB_pct99_18", "BB_pct99_19", 
            "BB_pct99_20", "BB_pct99_21", "BB_pct99_22", "BB_pct99_23", "TO_pct99_07", "TO_pct99_08", "TO_pct99_10", 
            "TO_pct99_12", "TO_pct99_13", "TO_pct99_14", "TO_pct99_16", "TO_pct99_18", "TO_pct99_19", "TO_pct99_20", 
            "TO_pct99_21", "TO_pct99_22", "TO_pct99_23")

## Read in rda files and assign to object
for (ii in pclist) {
  dta <- readRDS(paste0('data/', ii, '.rda'))

  assign(ii, dta)
}

pct01_07 <- cbind(BB_pct01_07, TO_pct01_07) %>%
  select(., -3)
pct01_08 <- cbind(BB_pct01_08, TO_pct01_08) %>%
  select(., -3)
pct01_10 <- cbind(BB_pct01_10, TO_pct01_10) %>%
  select(., -3)
pct01_12 <- cbind(BB_pct01_12, TO_pct01_12) %>%
  select(., -3)
pct01_13 <- cbind(BB_pct01_13, TO_pct01_13) %>%
  select(., -3)
pct01_14 <- cbind(BB_pct01_14, TO_pct01_14) %>%
  select(., -3)
pct01_16 <- cbind(BB_pct01_16, TO_pct01_16) %>%
  select(., -3)
pct01_18 <- cbind(BB_pct01_18, TO_pct01_18) %>%
  select(., -3)
pct01_19 <- cbind(BB_pct01_19, TO_pct01_19) %>%
  select(., -3)
pct01_20 <- cbind(BB_pct01_20, TO_pct01_20) %>%
  select(., -3)
pct01_21 <- cbind(BB_pct01_21, TO_pct01_21) %>%
  select(., -3)
pct01_22 <- cbind(BB_pct01_22, TO_pct01_22) %>%
  select(., -3)
pct01_23 <- cbind(BB_pct01_23, TO_pct01_23) %>%
  select(., -3)
```

Bind ALL DRIFTS into one dataframe
```{r}
allDrifts <- rbind(fulltrack07, fulltrack08, fulltrack10, 
                   fulltrack12, fulltrack13, fulltrack14, 
                   fulltrack16, fulltrack18, fulltrack19, 
                   fulltrack20, fulltrack21, fulltrack22, fulltrack23)
```

Where are loudest sound levels?
```{r}
pct01_07 %>%
  summarise_if(is.numeric, list(~ min(., na.rm=TRUE)))

```

Where are the quietest sound levels by band?
```{r}
pctlist <- c("07", "08", "10", "12", "14", "16", "18", "19",
             "20", "21", "22", "23")
# Doesn't show time stamp!
for (pc in pctlist) {
  data <- get(paste0("pct01_", pc)) %>%
    summarise_if(is.numeric, list(~ min(., na.rm=TRUE)))

  assign(paste0("min_", pc), data)
}

```

Where do we hear the most whale detections?
```{r}

```

Summary of each species: highest/lowest sound levels, number of detections
```{r}
allDrifts %>%

```