### 2023-08-08 02_AIS.Rmd
Update 02_AIS.Rmd
added Taiki's workflow to this

---
title: "AIS"
author: "Marina Bozinovic"
date: "2023-04-11"
output: html_document
---

#Introduction
AIS download from Marine Cadastre and match to buoy datasets.

Required Packages
```{r}
library(PAMscapes)
library(tidyverse)
```

```{r}
# Function
# Speeding up st_intersection (from https://github.com/r-spatial/sf/issues/801)

st_intersection_faster <- function(x,y,...){
#faster replacement for st_intersection(x, y,...)

  y_subset <-
    st_intersects(x, y) %>%
    unlist() %>%
    unique() %>%
    sort() %>%
    {y[.,]}

  st_intersection(x, y_subset,...)
}

```

```{r}
#########
# AIS
#########

##### Need to add more AIS dates to D: DRIVE!!!
AISlist <- list.files(path = paste0("data-raw/"), pattern = "AIS_", recursive = TRUE)

for (v in AISlist) {
  Adata <- read_csv(paste0("data-raw/", v), show_col_types = FALSE) %>%
    filter(Length >= 100) %>%                  # Filter out vessels over 100m in length
    dplyr::select(-SOG,-COG, -CallSign, -IMO, -Cargo, -Draft, -Status) %>%  #Remove unnecessary columns
    mutate(variable = paste0(substr(v, 16, 17), substr(v, 6, 7)))

  assign(paste0("AIS_", substr(v, 5, 10)), Adata)
}

AIS_sf <- AIS_091218 %>%
  st_as_sf(coords = c("LON", "LAT"), crs=4326, remove=F)
```

```{r}
##################################
# Create 10 km buffer around buoys (taken from M. McKenna's code)
# Should be malleable for each buoy's number to easily plug in.
##################################
tracks <- readRDS("data-raw/tracks.rda")

trk8 <- tracks %>% 
  dplyr::filter(station == "8") %>%
  st_as_sf(coords= c("Longitude", "Latitude"), crs=4326, remove=F)

pois <- st_transform(trk8, crs = 4326)    # Specify a crs
pts_buff <- st_buffer(pois, dist = 10000)  # Creates 10 km buffer around buoy points 10 km (10000m)
buff <- st_transform(pts_buff, crs = 4326)  # Specify crs for new buffer

### Crop full AIS to 10km buffer around drift 8 ###
AIS_crop <- st_crop(x = AIS_sf, y = buff)
```

```{r}
# Intersect buffer with AIS. Starting this on first 10 rows to see how slowly it goes. 
## FIX ME
#inters <- st_intersection(buff, AIS_crop[1:140,])

#Another way to view AIS points that fall within buffer. DOES NOT consider time.
AIS_within_buff <- AIS_crop[buff,]
f <- buff[AIS_within_buff,]
by <- join_by(closest(UTC >= BaseDateTime))

d <- full_join(buff, as.data.frame(AIS_within_buff), by)

e <- d[!duplicated(d[c('BaseDateTime')]),]

trk_AIS <- left_join(as.data.frame(trk8), as.data.frame(AIS_within_buff), by)

proj4string <- "+proj=longlat +datum=WGS84 +no_defs +ellps=WGS84 +towgs84=0,0,0"
trk_AIS_sf <- trk_AIS %>% st_as_sf(coords = c("Longitude", "Latitude"), crs=projcrs, remove=F)# %>%
   st_set_crs(4326)
trk_AIS_sf <- st_transform(trk_AIS_sf, crs = 4326)

#inters <- st_intersection_faster(buff, AIS_crop[1:140,])
#inters2 <- st_intersection_faster(buff, AIS_crop[141:200,])
#inters3 <- st_intersection_faster(buff, AIS_crop[201:300,])

#Another way to view AIS points that fall within buffer. DOES NOT consider time.

st_is_within_distance(AIS_crop, pois, dist = 10000)
st_distance(AIS_crop, pois, dist_fun = 10000)

st_set_geometry(AIS_crop, value = NULL)
st_set_geometry(pois, value = NULL)

AIS_crop <- as_Spatial(AIS_crop)
pois <- as_Spatial(pois)
AIS_within_buff <- as_Spatial(AIS_within_buff)
z <- sp_distance(AIS_crop, pois, features = TRUE, unit = "km")

data_places <- pois@data
data_places$distance <- sp_distance(AIS_within_buff, pois, features = TRUE, unit = "km")
View(data_places)

rgeos::gWithinDistance(AIS_crop, pois, features = TRUE, dist = 10000)
```
```{r}
# For each row in AIS df,
  # calculate distance to closest drift
  # return smallest distance
  # attach column to df

# 
ship_dist <- (geosphere::distHaversine(AIS_within_buff@coords[10:13,], pois))/1000
cbind(AIS_within_buff@data[10:13,], ship_dist)
```

-----------------------
From Taiki's AIS Workflow

```{r}
marcadFiles07 <- downloadMarCadAIS(tracks_SS_07, outDir = "D:/AIS", overwrite = FALSE, unzip = TRUE)
```

Subset to US West Coast
```{r}
subsetMarCadAIS(inDir="D:/AIS", outDir="D:/AIS/AIS_West", name='West_', 
                latRange=c(20,50), lonRange=c(-140,-110),
                overwrite=FALSE, progress=TRUE)
```
Now read in only ais points that come in contact with buoys.
```{r}
ais_near_07 <- readLocalAIS(tracks_SS_07, aisDir="D:/AIS/AIS_West", distance=10e3) %>%
   filter(vesselLength >= 100)  #remove vessels under 100m in length
View(ais_near_07)

gpsClose <- addAIS(tracks_SS_07, ais_near_07, interpType = 'close', interpTime = 120)
gpsNone <- addAIS(tracks_SS_07, ais_near_07, interpType = 'none')
gpsAll <- addAIS(tracks_SS_07, ais_near_07, interpType = 'all', interpTime = 120)

## Using PRE-ROUNDED dataset to try out
ais_near_raw7 <- readLocalAIS(tracks7, aisDir = "D:/AIS/AIS_West", distance=10e3) %>%
   filter(vesselLength >= 100)  #remove vessels under 100m in length
gpsClose_raw7 <- addAIS(tracks7, ais_near_raw7, interpType = 'close', interpTime = 120)
```

Plot
```{r}
ggplot() +
    geom_path(data=tracks_SS_07, aes(x=Longitude, y=Latitude), col='black') +
    geom_path(data=ais_near_07, aes(x=Longitude, y=Latitude, group=factor(MMSI), col=inDist)) +
    xlim(min(tracks_SS_07$Longitude), max(tracks_SS_07$Longitude)) +
    ylim(min(tracks_SS_07$Latitude), max(tracks_SS_07$Latitude))

baseGps <- ggplot(data=tracks7, aes(x=Longitude, y=Latitude), col='black') +
    geom_path() +
    xlim(min(tracks7$Longitude), max(tracks7$Longitude)) +
    ylim(min(tracks7$Latitude), max(tracks7$Latitude))

baseGps + geom_path(data=gpsClose_raw7, aes(x=shipLong, y=shipLat, group=factor(MMSI), color=shipDist < 10e3)) +
     geom_point(data=gpsClose_raw7, aes(x=shipLong, y=shipLat, color=shipDist < 10e3), size = 0.1) +
    ggtitle('CloseInterp')

# Viewing only ships within 10 km of buoys.
close2ships7 <- gpsClose %>% filter(shipDist < 10e3)
```