### 2023-08-08 manuscript.rmd
Update manuscript.rmd
added to BW and SW section of intro

---
title: "manuscript"
format: html
editor: visual
bibliography: references.bib
---

# Manuscript Draft

## Introduction

### Soundscapes

Sound is a critical form of communication for many species in the ocean. Since sound travels over four times faster in water than in air and can carry information over far distances @urick1983 , the use of sound has evolved as a critical component of communication, navigation, and foraging for many invertebrates, fishes and cetaceans @duarte2021 . Under the right conditions, sounds have the ability to travel across ocean basins @miksis-olds.

The total collection of sounds in an environment is called a soundscape, and these sources of sound are grouped into three broad categories: biophonies, sound produced by animals like cetaceans, fishes, and invertebrates; geophonies, natural sounds like rain, wind, and earthquakes; and anthropophonies, human-made sounds from shipping vessels, sonar, and survey activities (@miksis-olds, @duarte2021). Every region of the ocean experiences high spatio-temporal variation in sound levels and sources based on traits such as properties of water, organism diversity, human activities, and bathymetry @andré2011. Because of their unique sound signatures, ocean soundscapes are dynamic and can be characterized, measured, and monitored as indicators of ecosystem health or species diversity @radford2010 @haver2019 @weiss2021 @pijanowski2011 .

**\[what do they have to lose in a noisy environment?\]**

When noise levels reach above normal levels, organisms can be negatively impacted. Invertebrates and fishes may have stunted development and elevated stress responses (cite), fishes \[add negative behavior here\], and cetaceans' behavior, calling patterns, physiology and stress levels, ability to locate prey or conspecifics are impacted (cite). Each species is affected differently, some more sensitive to certain types of noise. While terrestrial ecosystem-level effects have been documented, (Buxton 2017, Shannon 2016, Francis 2012), those within aquatic ecosystems lack studies @kunc2016.

**\[Noise pollution is a growing concern and increasing in the ocean.\]**

Ocean noise pollution is among the anthropogenic changes that have recently grown to concerning levels @tyack2008, @hildebrand. The global ambient noise in the ocean is estimated to have increased by 12 decibels in the last few decades and has a close correlation to global economic trends @frisk2012, with commercial shipping accounting for most of the low-frequency anthropogenic noise in the ocean @hildebrand2009. Ships have been increasing in terms of gross tonnage and fleet size @hildebrand2009 (and age?), and now account for over 80% of the global trade, with an estimated 1.4% expected annual increase in shipping through 2027 @unctad.

These noise-generating activities are categorized as either incidental or deliberate @chou2021, @chahouri2022. Deliberate activities such as active sonar pingers, resource extraction from the seabed, seismic surveys, and echosounders are intentionally produce sound to measure an area and can utilize both high and low frequencies. Incidental activities like dredging, offshore development, and shipping produce sound as a by-product @hawkins2017. Anthropogenic noise can further be characterized by the frequency and duration of the sound. For example, shipping produces a continuous, low-frequency (\<200 Hz) sound that is widely distributed across the world's oceans @hildebrand2009. ~~Because this type of noise can travel great distances, they can have effects further away from the source~~. Sonar for military surveillance or research such as seafloor mapping occupies the mid-frequency bands, but attenuates quicker, therefore impacting a smaller acoustic space.

It remains unknown if chronic effects from continuous sound or acute effects from impulsive sounds can alter the normal behavior or physiology of cetaceans enough to affect reproduction and population levels. (though see @new2013 for habitat displacement affecting reproduction or survival)

Lower frequencies travel further, and higher frequencies tend to attenuate (@richardson2013 ), so the focus should be on local effects for that (@hildebrand2009 ). Ambient noise in shallow water has more variability than deeper water (@richardson2013). Climate change is also expected to alter the temperature, salinity, and pH of the ocean, causing an additional increase in global sound speed due to changing absorption qualities @affatati2022 @kunc2016 .

**\[This is an issue, but the extent of damage remains unclear, so it's difficult to produce legislation to that effect\]**

As soniferous animals, cetaceans have varying levels of sensitivity and vulnerability to certain types of noise. Mysticetes are more affected by low-frequency noise produced by large shipping vessels. Odontocetes produce and hear mid- to high-frequency sounds, so human activities that produce sounds at these frequencies interfere with these signals. The lines are not always so clear cut as odontocetes may be affected by the frequencies of motorized shipping typically known to affect baleen whales @aguilarsoto2006.

While many studies document negative responses of cetaceans to noise pollution with general frequency bands (cite), the circumstances surrounding these acoustic disturbances (chronic or impulsive, low, mid, or high frequency, shallow or deep water, overlap with anthropogenic activity) and their long-term effects are unclear.

habitat degradation

### Beaked Whales/Sperm Whales

[Beaked whales]{.underline} consist of 24 species within the family *Ziphiidae* and are among the least understood cetaceans and deepest divers in the ocean. They regularly dive for over an hour to depths greater than 800 m, and have been known to dive over three hours at a time to depths greater than 2000 m, followed by only a few minutes on the surface @tyack @baird2006. Their preference for deep water, coupled with a low-profile body makes them nearly invisible on the surface, so very little is known about them with visual observations @awbery2022 . Much of what is known comes from strandings and stomach content analysis @west2017. Even when sighting events occur, they risk misidentification due to the paucity of data regarding morphological variations in some species @aguilardesoto2017. Tagging studies prove useful, but difficult to scale/acquire sizable dataset @baird2006 @baird2008 @tyack .

Beaked whales are also known to be unusually sensitive to noise. There are several documented cases of naval mid-frequency active sonar (MFAS) activities coinciding with mass stranding events by beaked whales (see @simonis2020, table 1 in @weilgart, @deruiter2013 , @stanistreet2022, @bernaldodequiros, @fernández2005 ). This occurs for two reasons: 1) their anti-predation adaptation to avoid killer whales (*Orcinus orca*)(synchronized foraging, followed by silent, scattered vertical ascent) mimics that of anthropogenic noise and leads to strandings @aguilardesoto2020; 2) the same physiological adaptations that allow for extreme diving are observed to be deadly when forced to alter behavior and flee the noise source more so than other odontocetes @tyack @cox2006 @joyce2020 . Furthermore, these responses vary based on the individuals' age, sex, past experience with noise, or current activity @weilgart.

Despite their visually cryptic nature, their sound signals are very well-described, down to the species level @baumann-pickering2013, @baumann-pickering2014. Beaked whale produce clicks, pulses, and buzzes during foraging starting at the 400(?) m range @tyack. With increasing acoustic surveys in the literature, new vocalization patterns are being detected globally @manzano-roth2023 and notable temporal patterns are emerging @yamada2019.

Sperm whales (*Physeter macrocephalus*) also occupy deep ocean habitats, but their dive depths and times are reduced in comparison. Sperm whales have been known to dive regularly between 400-600 m and often over 1000 m @watkins1993 @awbery2022 @amano2003. They also appear to have more acoustic tolerance than beaked whales @miller2009, @miller2022. While avoidance responses in sperm whales have been recorded following naval sonar events @curé2013 @stanistreet2022 @sivle2012 , the full extent is unknown.

Sperm whales also emit series of targeted clicks when socializing (@marcoux2006 @fais2015) and when foraging @fais2016, but it's unlikely they debilitate their prey with their vocalizations as previously hypothesized @norris1972 . More is known about sperm whales than beaked whales as they are less visually cryptic and have a

Species studied: \
- ZC Cuvier's beaked whale *Ziphius cavirostris (IUCN status: least concern)\
-* BW43, Perrin's beaked whale *M*. *perrini (IUCN status: endangered)* confirmed in 2021 @baumann-pickering2013 @barlow2021\
- BW37V, possibly Hubbs' beaked whale *M. carlhubbsi (IUCN status: data deficient)* @simonis2020*\
**-* BB, --???\
- BWC, possibly ginkgo-toothed beaked whale *M. ginkgodens (IUCN status: data deficient)* @simonis2020*\
**-* PM, sperm whale *(IUCN status: vulnerable)*

**Effects of Acoustic Disturbance**

The specific effects of acoustic interference from human activity in beaked and sperm whales are well-documented. Intense noise events can affect behavior through habitat displacement, spending time and effort fleeing instead of foraging, communication and echolocation click disruption, increasing call intensity (cite?) (@deruiter2013, @aguilarsoto2006 , @stanistreet2022, @joyce2020, @falcone2017, @cholewiak2017 ), cause physiological harm like hearing loss, tissue damage, and elevated stress levels (see @hooker2019) even at far distances (up to 100 km away @falcone2017), masksignals for communication (@erbe2014), environmental cues (@duarte2021), or mating @weilgart.\
- see Cox et al 2006 for impacts of sound on BW

Under the Marine Mammal Protection Act, all these species are federally protected, several receive addition protection under the Endangered Species Act due to a vulnerable or endangered conservation status @taylor @perrin , but some remain listed as "data deficient" ( @pitman @ginkgodens). The latter designation has been scrutinized for not illiciting the same conservation urgency as species whose populations are well-documented therefore gets lowered on the priority list (@parsons2016).

### Passive Acoustic Monitoring

Passive acoustic monitoring is a tool that allows us to record and listen to the soundscape. It is not fit for all applications or ecological questions, but beaked and sperm whales are ideal study candidates for this type of monitoring because they are often better heard than seen ( @fleishman ). This method provides reliable estimates @arranz2023.

Bottom-mounted/static: high temporal resolution, low spatial resolution @mellinger

Acoustic tags:

Towed arrays:

Tradeoffs include costly equipment or deployment, data/battery storage requirements, chance of return, spatial coverage, temporal coverage, capturing fine-scale movements, number of individuals, parsing out individuals, through inclement weather or night-time @johnson2009 @fleishman

One type of passive acoustic recorder called a drifting buoy that drifts freely with the currents for days or weeks at a time. It allows us to capture the dynamic nature of sound over a large spatial area and especially in the deep, offshore, and less studied waters. Buoys offer greater spatial and temporal coverage compared to bottom-mounted or stationary hydrophones, ideal for detection of offshore or mobile species. Studying beaked whales can be challenging due to how offshore they live (@robbins2022). The goal with passive acoustic monitoring is to improve current species distribution models of beaked and sperm whales that utilize visual survey methods only. Historically, it has been challenging to record foraging calls at several hundred meters of depth @barlow2005.

Its basic components consist of a satellite transmitter above water to monitor GPS information and an acoustic recorder attached by line at \~100m of depth. In fact, most of what is known about these deep divers is because of acoustic monitoring. Similar designs of drifting buoys have been used in estuarine and coral reef soundscape studies @lillis2018.

### **AIS**

Automatic Identification System. Ships are abundant throughout the study area due to major US ports. This was collected to investigate patterns of co-occurrence between vessels and whales and sound levels in those moments.

@aguilarsoto2006 -- elevated sound levels during close encounters between ships and beaked whales can be detrimental.

How have other studies separated vessel noise (contribution to soundscape) from ambient soundscape? See @weiss2021 and @haver2020. Power spectral density ambient sound levels.

**Laws and Regulations** \[Where and to what extent is noise pollution regulated?\]

see @chou2021 for a list of documents, reports, programs, efforts in ocean noise reduction

Marine Strategy Framework Directive by European Union establishes a directive about environmental regulation on marine noise, specifically requiring member states to adopt strategies in compliance with threshold values for impulsive and continuous noise. @directiv2008, @borsanij.f.2023 Within their Marine Strategy Framework Directive, the European Union is the only entity to date that has recognized noise as a form of pollution and restricts noise from reaching levels that may cause negative effects to the environment @directiv2008. [What about UK Marine Strategy?] See @chou2021 for larger list of efforts.

For mid-frequency cetaceans, auditory impairment threshold levels are as follows:\
peak sound pressure level (Impulsive) - 230 dB, recommended accumulation period of 24 hrs (impulsive) - 185 dB, recommended accumulation period of 24 hrs (impulsive) - 198 dB (@noaa2018) .

IMO guidelines from 2014 are being updated by July 2023. @internationalmaritimeorganization

Canada is developing an Ocean Noise Strategy

In the US, it's difficult to achieve compliance from Navy.

### Research Objective

\
Shipping and offshore activities are expected to increase in the future, causing elevated oceanic sound levels. This is a form of habitat degradation, so there is a need to understand how human activities change and contribute to the overall soundscape in space and time. It's a challenge to acquire information on these federally protected and acoustically-sensitive species but becoming increasingly important to understand where habitats overlap with human activity. Acoustics can serve as important indicators of ecosystem health, but also of how humans are changing the environment.\
\
Characterize the acoustic habitat of beaked and sperm whales in the California Current Ecosystem using soundscape metrics and vessel tracking. Characterizing the acoustic environment and identifying co-occurrence of whales and anthropogenic sound has been done in a similar study by @weiss2021. This serves as baseline monitoring of soundscape for offshore areas.

**\[Gaps in literature, rationale for research\]**

"To date, studies have focused on this and this, but not on this".This work will contribute to establishing baseline conditions of noise exposure to inform wind farm development and shipping lane changes. \[add here\]

It'll also add to the knowledge base of beaked and sperm whale distribution through acoustic detections.

Both of these outcomes are of high importance to managers and agencies especially NOAA and BOEM (see 2023-2024 report). Within their Ocean Noise Strategy Roadmap, NOAA is specifically seeking 1) a quantification of the spatial and temporal variability of ambient noise conditions, and 2) understanding of how anthropogenic sound sources contribute to the soundscape. Co-occurrence of buoys and vessels will provide fine-scale snapshots of sound levels surrounding vessels and concurrent whale foraging activity. With AIS, we can calculate distance from buoy to vessel and sound levels in those moments. And 3) NOAA is seeking more information on vocally active species' distributions. This is part of a bill introduced by Congress to "assess underwater sound in high-priority" environments (https://www.congress.gov/bill/117th-congress/house-bill/6987).

\[add here\]

Quantifying soundscapes in a standardized way and monitoring these changes over time are priorities for marine managers and researchers right now. Connecting human noise sources to whale habitat in space and time is needed for efficient management @haver2019.

"Increased understanding of temporal and spatial distribution of beaked whales will enable more effective planning for noise-generating activities to occur where and when fewer sensitive animals are present." @hooker2019

"Very large-scale studies, conducted on the spatial, temporal and spectral scales at which these highly mobile predators live their lives, may be needed to evaluate the influence of changes in a marine mammal's acoustic environment." @erbe2014

"Research should be conducted on how the level of anthropogenic activity (such as the types and numbers of vessels) is related to resulting noise levels." @hildebrand2009

## Methods

### **Study Area**

\
The California Current Ecosystem extends along the US' Economic Exclusion Zone along the west coast. It is characterized by... It's a highly productive and complex ecosystem that contains five current and one proposed national marine sanctuaries \[add National Marine Sanctuary Act support\], supports several wild fisheries, and is home to many protected species. Because this area has high economic value and natural ecological activity, it's a well-managed and well-researched area. "Ambient noise off the coast of California has increased many-fold over the past several decades..." (@moore2013 + other sources w/in that).

In 2021, California, Washington, and Oregon ranked 3rd, 5th, and 23rd in most trafficked U.S. waterways, respectively (@unitedstatesarmycorpsofengineersusace2018. Lots of overlap between vessel activity from busy ports and important habitat for protected species.

\[Can discuss the warm and cool phases that occurred during study period.\]

see @harvey for report on conditions

### Data Collection and Processing

**\[Overall data collection efforts\]**\
Acoustic detections were collected with the use of free-drifting Drifting Acoustic Spar Buoy Recorders (hereafter buoys). These were deployed as part of the California Current Ecosystem Survey (CCES) spanning the offshore waters from southern Canada to northern Mexico in summer of 2018. These buoys contained two satellite geo-location trackers, two hydrophones, etc and collected a total of 1910 hours of data. SWFSC ran acoustic detections in PAMGuard. Eight buoys were lost at sea and therefore removed from the analysis. These contain 1, 5, 10, 25, 75, 95, and 99 percentile distribution from each metric.

**\[Add details about acoustic data collection information.\]**\
Soundscape metrics were duty-cycled (most 18min off, 2 minutes on, but some variation in "off" times) and later processed by the NOAA team with calls and clicks identified to the species level (see @simonis2020 for details on this survey and @rand2022 for duty-cycle study). Soundscape metrics were binned into low, medium, high frequency bands. Can include software used, sampling rate?

**\[Post-processing details, data cleaning\]**\
Soundscape metrics were rounded to the nearest 20 minutes (:00, :20, :40) and converted to UTC. These include third-octave level bands and broadband (20-24,000 Hz) metrics. ~~One way to measure sound is with a metric called a third octave level, which splits the sound spectrum into many bands.~~ The band at 125 Hertz is often associated with shipping vessel noise and the 2-5 kHz range with mid-frequency naval sonar ( @garrett2016 ). Beaked whales are known to be sensitive to sonar in this latter range (cite). Several TOL levels were chosen as proxies to vessel or sonar frequencies, ranging from TOL 63 to 20,000.

Whale detections were rounded to nearest 20 minutes. A separate column was created as a common join field that contained nearest timestamp in UTC to that of drifter tracks. Two types of BW species (?BW and ?BWC) were removed from analysis as the species is unknown and unsure. Whale detections were reduced to "presence/absence" from number of clicks, estimated individuals, etc.

Drifter tracks were rounded to nearest 20 minutes, lost buoys (1,2,3,5,6,9,11) and corrupted buoys (4,17) were removed. Drift 14 and 15 are the same. These three major datasets were joined together based on a common buoy number (column "station") and time field column "UTC").

**\[Environmental Data\]**\
Environmetal and oceanographic variables thought to be predictors of BW/SW habitat were extracted from NOAA's Environmental Research Division Data Access Program (ERDDAP) and Hybrid Coordinate Ocean Model (HYCOM) using matchEnvData function in package PAMmisc. Variables at depth were calculated using function from Taiki and formula from @mccullough2021. Distance to slope and bathymetric slope were calculated from a bathymetry TIF taken from GEBCO.

\[Can show table of variables and justification by papers (Virgili et al. 2022)\]

use @torreblanca2022 for deep habitat info\
see @johnston2008 and @manzano-roth2023 for seamount habitat importance

| Covariate                     | Unit | Source                                                          | Justification                                                                                                         |
|-----------------|-----------------|-----------------|---------------------|
|                               |      | \[can use A, B, C to avoid repetition, cite in caption\]        |                                                                                                                       |
| Distance to continental slope |      | General Bathymetric Chart of the Oceans (GEBCO) @weatherall2015 |                                                                                                                       |
| Bathymetric slope             |      | GEBCO tif                                                       |                                                                                                                       |
|                               |      |                                                                 |                                                                                                                       |
| Wind curl stress              |      | ERDDAP                                                          | "Positive wind stress curl lifts the thermocline, which results in enhanced primary production" (Manocci et al 2015). |
|                               |      |                                                                 |                                                                                                                       |
|                               |      |                                                                 |                                                                                                                       |
|                               |      |                                                                 |                                                                                                                       |
|                               |      |                                                                 |                                                                                                                       |
|                               |      |                                                                 |                                                                                                                       |
|                               |      |                                                                 |                                                                                                                       |
|                               |      |                                                                 |                                                                                                                       |
|                               |      |                                                                 |                                                                                                                       |
|                               |      |                                                                 |                                                                                                                       |

**\[AIS\]**\
Automatic Identification System vessel tracking data were collected from Marine Cadastre in order to evaluate the effect of spatial proximity to the buoys. This step is important in understanding how ship characteristics contribute to various frequency bands. I filtered out any vessels under 100 meters in length in order to focus only on large vessels. By exploring buoy/vessel co-occurrence with a 10-km "listening space" around the buoy (McKenna) can allow us to investigate patterns of when whales are calling and the vessel activity during co-occurrence.

AIS was downloaded, processed (converted to UTC), and matched to each drifter using Taiki's PAMscapes package and functions.

**\[Big analysis: statistics, modeling details\]**

Use Random Forest model to identify most important predictors.

Can use non-parametric Kruskal-Wallis one-way analysis of variance test to examine *if number of whale detections per hour differed depending on presence/absence of vessels* (see Trickey et al. 2022).

[Example of GAM]{.underline}: @mannocci2015

AIC explanation @symonds2011

## Results

[Describe simple statistical relationships, correlations, exploratory data findings (highest, lowest, etc., time of day ( @ziegenhorn2023 ) , drift with most detections/day)]

\[Weekly/daily plot of acoustic presence of whales and ships over time from each drift? See JS Trickey et al. 2022, Fig 3\]

Look at power spectral density? @haver2020

## Discussion

**\[Known limitations to this study\]\
**Duty cycle whale detection and soundscape method knowingly omits 18 of 20 minutes of data. A sperm whale's true location relative to buoy is unknown so preferred habitat may not reflect the environmental variables that are associated with buoy that detects it (Barkley et al. 2021) Rounding tracks and soundscape metrics date/times to nearest 20 minutes is not perfect. Often, times did not line up with :00, :20, :40 and were either omitted or duplicated.

This study only captures half of the year. If BW/SW have seasonal preferences outside of the study period, it was not captured here.

Are BW or SW known to have seasonal movements? @robbins2022 suggest they do, though there's conflicting evidence.

Non-AIS vessels or small vessels (fishing, recreational) may contribute to the soundscape in relevant frequency bands (@hermannsen2019 , @hildebrand2009 ) and this does not get captured here.

\[Add here.\]

**\[Recommendations?\]**

## Acknowledgements

Taiki - help with code!\
Megan McKenna - guidance on AIS questions.

\[Add here.\]