### 2023-09-13 07_summary_stats.Rmd
Update 07_summary_stats.Rmd
removed broadband metrics

---
title: "07_summary_stats"
author: "Marina Bozinovic"
date: "2023-06-28"
output: html_document
---

```{r setup, include=FALSE}
library(tidyverse)
library(lubridate)
library(PAMpal)
library(PAMmisc)
library(PAMscapes)
library(patchwork)
library(knitr)
library(kableExtra)
```

Read in RDS files of each buoy + all environ variables + AIS (pending)
```{r}
ftlist <- dir("data/", recursive=TRUE, full.names=TRUE, pattern="fulltrack")

for (ft in ftlist) {
  df <- readRDS(ft)
  assign(paste0("fulltrack", substr(ft, 15, 16)), df)
}
```

Read in RDA files for percentiles
```{r}
## Make list of rda files to read in
pclist <- c("pct10_07", "pct10_08", "pct10_10", "pct10_12", "pct10_13",
            "pct10_14", "pct10_16", "pct10_18", "pct10_19", "pct10_20",
            "pct10_21", "pct10_22", "pct10_23", "pct90_07", "pct90_08",
            "pct90_10", "pct90_12", "pct90_13", "pct90_14", "pct90_16",
            "pct90_18", "pct90_19", "pct90_20", "pct90_21", "pct90_22",
            "pct90_23", "pct50_07", "pct50_08", "pct50_10", "pct50_12",
            "pct50_13", "pct50_14", "pct50_16", "pct50_18", "pct50_19", 
            "pct50_20", "pct50_21", "pct50_22", "pct50_23")

## Read in rda files and assign to object
for (ii in pclist) {
  dta <- readRDS(paste0('data/', ii, '.rda'))

  assign(ii, dta)
}

```

Create df from 1 percentile
```{r}
# 1 Percentile
pct01_07 <- cbind(BB_pct01_07, TO_pct01_07) %>%
  select(., -3)
pct01_08 <- cbind(BB_pct01_08, TO_pct01_08) %>%
  select(., -3)
pct01_10 <- cbind(BB_pct01_10, TO_pct01_10) %>%
  select(., -3)
pct01_12 <- cbind(BB_pct01_12, TO_pct01_12) %>%
  select(., -3)
pct01_13 <- cbind(BB_pct01_13, TO_pct01_13) %>%
  select(., -3)
pct01_14 <- cbind(BB_pct01_14, TO_pct01_14) %>%
  select(., -3)
pct01_16 <- cbind(BB_pct01_16, TO_pct01_16) %>%
  select(., -3)
pct01_18 <- cbind(BB_pct01_18, TO_pct01_18) %>%
  select(., -3)
pct01_19 <- cbind(BB_pct01_19, TO_pct01_19) %>%
  select(., -3)
pct01_20 <- cbind(BB_pct01_20, TO_pct01_20) %>%
  select(., -3)
pct01_21 <- cbind(BB_pct01_21, TO_pct01_21) %>%
  select(., -3)
pct01_22 <- cbind(BB_pct01_22, TO_pct01_22) %>%
  select(., -3)
pct01_23 <- cbind(BB_pct01_23, TO_pct01_23) %>%
  select(., -3)

pct01_07$buoy <- ("07")
pct01_08$buoy <- ("08")
pct01_10$buoy <- ("10")
pct01_12$buoy <- ("12")
pct01_13$buoy <- ("13")
pct01_14$buoy <- ("14")
pct01_16$buoy <- ("16")
pct01_18$buoy <- ("18")
pct01_19$buoy <- ("19")
pct01_20$buoy <- ("20")
pct01_21$buoy <- ("21")
pct01_22$buoy <- ("22")
pct01_23$buoy <- ("23")

pct01_all <- rbind(pct01_07, pct01_08, pct01_10, pct01_12, pct01_13,
                   pct01_14, pct01_16, pct01_18, pct01_19, pct01_20,
                   pct01_21, pct01_22, pct01_23)

# Group by and summarize
pct01_all %>%
  group_by(buoy) %>%
  summarize(max(`TOL_125`))

```

Create df from 99 percentile
```{r}
# 99 Percentile
pct99_07 <- cbind(BB_pct99_07, TO_pct99_07) %>%
  select(., -3)
pct99_08 <- cbind(BB_pct99_08, TO_pct99_08) %>%
  select(., -3)
pct99_10 <- cbind(BB_pct99_10, TO_pct99_10) %>%
  select(., -3)
pct99_12 <- cbind(BB_pct99_12, TO_pct99_12) %>%
  select(., -3)
pct99_13 <- cbind(BB_pct99_13, TO_pct99_13) %>%
  select(., -3)
pct99_14 <- cbind(BB_pct99_14, TO_pct99_14) %>%
  select(., -3)
pct99_16 <- cbind(BB_pct99_16, TO_pct99_16) %>%
  select(., -3)
pct99_18 <- cbind(BB_pct99_18, TO_pct99_18) %>%
  select(., -3)
pct99_19 <- cbind(BB_pct99_19, TO_pct99_19) %>%
  select(., -3)
pct99_20 <- cbind(BB_pct99_20, TO_pct99_20) %>%
  select(., -3)
pct99_21 <- cbind(BB_pct99_21, TO_pct99_21) %>%
  select(., -3)
pct99_22 <- cbind(BB_pct99_22, TO_pct99_22) %>%
  select(., -3)
pct99_23 <- cbind(BB_pct99_23, TO_pct99_23) %>%
  select(., -3)

# Add column with buoy number
pct99_07$buoy <- ("07")
pct99_08$buoy <- ("08")
pct99_10$buoy <- ("10")
pct99_12$buoy <- ("12")
pct99_13$buoy <- ("13")
pct99_14$buoy <- ("14")
pct99_16$buoy <- ("16")
pct99_18$buoy <- ("18")
pct99_19$buoy <- ("19")
pct99_20$buoy <- ("20")
pct99_21$buoy <- ("21")
pct99_22$buoy <- ("22")
pct99_23$buoy <- ("23")

pct99_all <- rbind(pct99_07, pct99_08, pct99_10, pct99_12, pct99_13,
                   pct99_14, pct99_16, pct99_18, pct99_19, pct99_20,
                   pct99_21, pct99_22, pct99_23)
```

Plot 99pct and 1pct by buoy
```{r fig.height=5, fig.width=13}
# Group by and summarize
pct99_all %>%
  group_by(buoy) %>%
  summarize(max(`TOL_125`)) %>%
  ggplot(aes(x = buoy, y = `max(TOL_125)`)) + geom_bar(stat = "identity")

#TOL timeline by buoy
pct99_all %>%
  ggplot(aes(x = dateTime, y = TOL_2000, color = buoy)) + geom_line()

# Average dB by hour for each frequency
pct99_16 %>%
  group_by(time = floor_date(dateTime, "1 hour")) %>%
  summarize(mean_BB = mean(`BB_20-24000`),
            meanTOL_125 = mean(TOL_125),
            meanTOL_250 = mean(TOL_250),
            meanTOL_500 = mean(TOL_500),
            meanTOL_2000 = mean(TOL_2000),
            meanTOL_4000 = mean(TOL_4000),
            meanTOL_20000 = mean(TOL_20000)) %>%
  pivot_longer(cols = mean_BB:meanTOL_20000,
               names_to = "frequency",
               values_to="dB") %>%
  ggplot(aes(x = time, y = dB)) + geom_line(aes(color = frequency)) #+
  #geom_point(data = allDrifts, aes(color = species))
# Show more dates/times on x axis
# 99pct, buoy 16 has a strong peak later in journey at 4kHz

ggplot() +
  geom_point(data = pct01_23, aes(x = lubridate::round_date(dateTime, "2 hours"), y = TOL_125), color = 'steelblue') +
  #geom_line(data = pct99_23, aes(x = dateTime, y = TOL_125), color = 'coral') +
  facet_grid()

```

Bind percentile df (pctx_x) with GPS track df (GPSwhalex)

Buoy 7
```{r}
# Create columns with percentiles and BIND
pct50_07_select <- pct50_07 %>% select(UTC, TOL_125, TOL_2000) %>% 
  rename(pct50_TOL125 = TOL_125,
         pct50_TOL2000 = TOL_2000)
pct10_07_select <- pct10_07 %>% select(UTC, TOL_125, TOL_2000) %>%
   rename(pct10_TOL125 = TOL_125,
         pct10_TOL2000 = TOL_2000)
pct90_07_select <- pct90_07 %>% select(UTC, TOL_125, TOL_2000) %>%
   rename(pct90_TOL125 = TOL_125,
         pct90_TOL2000 = TOL_2000)

# Bind
pct07 <- cbind(pct10_07_select, pct50_07_select, pct90_07_select)

# Join GPS points with the three percentiles
pct07_whl <- timeJoin(GPSwhale7, pct07, thresh = 3600, interpolate = FALSE)

#Summarize
buoy7pctSummary <- pct07_whl %>%
  group_by(species) %>%
  summarize(pct10_125_mean = mean(pct10_TOL125),
            pct50_125_mean = mean(pct50_TOL125),
            pct90_125_mean = mean(pct90_TOL125),
            pct10_2000_mean = mean(pct10_TOL2000),
            pct50_2000_mean = mean(pct50_TOL2000),
            pct90_2000_mean = mean(pct90_TOL2000))
```

Buoy 8
```{r}
# Create columns with percentiles and BIND
pct50_08_select <- pct50_08 %>% select(UTC, TOL_125, TOL_2000) %>% 
  rename(pct50_TOL125 = TOL_125,
         pct50_TOL2000 = TOL_2000)
pct10_08_select <- pct10_08 %>% select(UTC, TOL_125, TOL_2000) %>%
   rename(pct10_TOL125 = TOL_125,
         pct10_TOL2000 = TOL_2000)
pct90_08_select <- pct90_08 %>% select(UTC, TOL_125, TOL_2000) %>%
   rename(pct90_TOL125 = TOL_125,
         pct90_TOL2000 = TOL_2000)

# Bind
pct08 <- cbind(pct10_08_select, pct50_08_select, pct90_08_select)

# Join GPS points with the three percentiles
pct08_whl <- timeJoin(GPSwhale8, pct08, thresh = 3600, interpolate = FALSE)

#Summarize
buoy8pctSummary <- pct08_whl %>%
  group_by(species) %>%
  summarize(pct10_125_mean = mean(pct10_TOL125),
            pct50_125_mean = mean(pct50_TOL125),
            pct90_125_mean = mean(pct90_TOL125),
            pct10_2000_mean = mean(pct10_TOL2000),
            pct50_2000_mean = mean(pct50_TOL2000),
            pct90_2000_mean = mean(pct90_TOL2000))
```

Buoy 10
```{r}
# Create columns with percentiles and BIND
pct50_10_select <- pct50_10 %>% select(UTC, TOL_125, TOL_2000) %>% 
  rename(pct50_TOL125 = TOL_125,
         pct50_TOL2000 = TOL_2000)
pct10_10_select <- pct10_10 %>% select(UTC, TOL_125, TOL_2000) %>%
   rename(pct10_TOL125 = TOL_125,
         pct10_TOL2000 = TOL_2000)
pct90_10_select <- pct90_10 %>% select(UTC, TOL_125, TOL_2000) %>%
   rename(pct90_TOL125 = TOL_125,
         pct90_TOL2000 = TOL_2000)

# Bind
pct10 <- cbind(pct10_10_select, pct50_10_select, pct90_10_select)

# Join GPS points with the three percentiles
pct10_whl <- timeJoin(GPSwhale10, pct10, thresh = 3600, interpolate = FALSE)

#Summarize
buoy10pctSummary <- pct10_whl %>%
  group_by(species) %>%
  summarize(pct10_125_mean = mean(pct10_TOL125),
            pct50_125_mean = mean(pct50_TOL125),
            pct90_125_mean = mean(pct90_TOL125),
            pct10_2000_mean = mean(pct10_TOL2000),
            pct50_2000_mean = mean(pct50_TOL2000),
            pct90_2000_mean = mean(pct90_TOL2000))
```

Buoy 12
```{r}
# Create columns with percentiles and BIND
pct50_12_select <- pct50_12 %>% select(UTC, TOL_125, TOL_2000) %>% 
  rename(pct50_TOL125 = TOL_125,
         pct50_TOL2000 = TOL_2000)
pct10_12_select <- pct10_12 %>% select(UTC, TOL_125, TOL_2000) %>%
   rename(pct10_TOL125 = TOL_125,
         pct10_TOL2000 = TOL_2000)
pct90_12_select <- pct90_12 %>% select(UTC, TOL_125, TOL_2000) %>%
   rename(pct90_TOL125 = TOL_125,
         pct90_TOL2000 = TOL_2000)

# Bind
pct12 <- cbind(pct10_12_select, pct50_12_select, pct90_12_select)

# Join GPS points with the three percentiles
pct12_whl <- timeJoin(GPSwhale12, pct12, thresh = 3600, interpolate = FALSE)

#Summarize
buoy12pctSummary <- pct12_whl %>%
  group_by(species) %>%
  summarize(pct10_125_mean = mean(pct10_TOL125),
            pct50_125_mean = mean(pct50_TOL125),
            pct90_125_mean = mean(pct90_TOL125),
            pct10_2000_mean = mean(pct10_TOL2000),
            pct50_2000_mean = mean(pct50_TOL2000),
            pct90_2000_mean = mean(pct90_TOL2000))
```

Buoy 13
```{r}
# Create columns with percentiles and BIND
pct50_13_select <- pct50_13 %>% select(UTC, TOL_125, TOL_2000) %>% 
  rename(pct50_TOL125 = TOL_125,
         pct50_TOL2000 = TOL_2000)
pct10_13_select <- pct10_13 %>% select(UTC, TOL_125, TOL_2000) %>%
   rename(pct10_TOL125 = TOL_125,
         pct10_TOL2000 = TOL_2000)
pct90_13_select <- pct90_13 %>% select(UTC, TOL_125, TOL_2000) %>%
   rename(pct90_TOL125 = TOL_125,
         pct90_TOL2000 = TOL_2000)

# Bind
pct13 <- cbind(pct10_13_select, pct50_13_select, pct90_13_select)

# Join GPS points with the three percentiles
pct13_whl <- timeJoin(GPSwhale13, pct13, thresh = 3600, interpolate = FALSE)

#Summarize
buoy13pctSummary <- pct13_whl %>%
  group_by(species) %>%
  summarize(pct10_125_mean = mean(pct10_TOL125),
            pct50_125_mean = mean(pct50_TOL125),
            pct90_125_mean = mean(pct90_TOL125),
            pct10_2000_mean = mean(pct10_TOL2000),
            pct50_2000_mean = mean(pct50_TOL2000),
            pct90_2000_mean = mean(pct90_TOL2000))
```

Buoy 14 # Issue binding pct50 with others due to differing number of rows
```{r}
# Create columns with percentiles and BIND
pct50_14_select <- pct50_14 %>% select(UTC, TOL_125, TOL_2000) %>% 
  rename(pct50_TOL125 = TOL_125,
         pct50_TOL2000 = TOL_2000)
pct10_14_select <- pct10_14 %>% select(UTC, TOL_125, TOL_2000) %>%
   rename(pct10_TOL125 = TOL_125,
         pct10_TOL2000 = TOL_2000)
pct90_14_select <- pct90_14 %>% select(UTC, TOL_125, TOL_2000) %>%
   rename(pct90_TOL125 = TOL_125,
         pct90_TOL2000 = TOL_2000)

# Bind
#pct14 <- cbind(pct10_14_select, pct50_14_select, pct90_14_select)

# Join GPS points with the three percentiles
pct14_whl <- timeJoin(GPSwhale14, pct14, thresh = 3600, interpolate = FALSE)

#Summarize
buoy14pctSummary <- pct14_whl %>%
  group_by(species) %>%
  summarize(pct10_125_mean = mean(pct10_TOL125),
            pct50_125_mean = mean(pct50_TOL125),
            pct90_125_mean = mean(pct90_TOL125),
            pct10_2000_mean = mean(pct10_TOL2000),
            pct50_2000_mean = mean(pct50_TOL2000),
            pct90_2000_mean = mean(pct90_TOL2000))
```
Buoy 16
```{r}
# Create columns with percentiles and BIND
pct50_16_select <- pct50_16 %>% select(UTC, TOL_125, TOL_2000) %>% 
  rename(pct50_TOL125 = TOL_125,
         pct50_TOL2000 = TOL_2000)
pct10_16_select <- pct10_16 %>% select(UTC, TOL_125, TOL_2000) %>%
   rename(pct10_TOL125 = TOL_125,
         pct10_TOL2000 = TOL_2000)
pct90_16_select <- pct90_16 %>% select(UTC, TOL_125, TOL_2000) %>%
   rename(pct90_TOL125 = TOL_125,
         pct90_TOL2000 = TOL_2000)

# Bind
pct16 <- cbind(pct10_16_select, pct50_16_select, pct90_16_select)

# Join GPS points with the three percentiles
pct16_whl <- timeJoin(GPSwhale16, pct16, thresh = 3600, interpolate = FALSE)

#Summarize
buoy16pctSummary <- pct16_whl %>%
  group_by(species) %>%
  summarize(pct10_125_mean = mean(pct10_TOL125),
            pct50_125_mean = mean(pct50_TOL125),
            pct90_125_mean = mean(pct90_TOL125),
            pct10_2000_mean = mean(pct10_TOL2000),
            pct50_2000_mean = mean(pct50_TOL2000),
            pct90_2000_mean = mean(pct90_TOL2000))
```

Buoy 18
```{r}
# Create columns with percentiles and BIND
pct50_18_select <- pct50_18 %>% select(UTC, TOL_125, TOL_2000) %>% 
  rename(pct50_TOL125 = TOL_125,
         pct50_TOL2000 = TOL_2000)
pct10_18_select <- pct10_18 %>% select(UTC, TOL_125, TOL_2000) %>%
   rename(pct10_TOL125 = TOL_125,
         pct10_TOL2000 = TOL_2000)
pct90_18_select <- pct90_18 %>% select(UTC, TOL_125, TOL_2000) %>%
   rename(pct90_TOL125 = TOL_125,
         pct90_TOL2000 = TOL_2000)

# Bind
pct18 <- cbind(pct10_18_select, pct50_18_select, pct90_18_select)

# Join GPS points with the three percentiles
pct18_whl <- timeJoin(GPSwhale18, pct18, thresh = 3600, interpolate = FALSE)

#Summarize
buoy18pctSummary <- pct18_whl %>%
  group_by(species) %>%
  summarize(pct10_125_mean = mean(pct10_TOL125),
            pct50_125_mean = mean(pct50_TOL125),
            pct90_125_mean = mean(pct90_TOL125),
            pct10_2000_mean = mean(pct10_TOL2000),
            pct50_2000_mean = mean(pct50_TOL2000),
            pct90_2000_mean = mean(pct90_TOL2000))
```

Buoy 19
```{r}
# Create columns with percentiles and BIND
pct50_19_select <- pct50_19 %>% select(UTC, TOL_125, TOL_2000) %>% 
  rename(pct50_TOL125 = TOL_125,
         pct50_TOL2000 = TOL_2000)
pct10_19_select <- pct10_19 %>% select(UTC, TOL_125, TOL_2000) %>%
   rename(pct10_TOL125 = TOL_125,
         pct10_TOL2000 = TOL_2000)
pct90_19_select <- pct90_19 %>% select(UTC, TOL_125, TOL_2000) %>%
   rename(pct90_TOL125 = TOL_125,
         pct90_TOL2000 = TOL_2000)

# Bind
pct19 <- cbind(pct10_19_select, pct50_19_select, pct90_19_select)

# Join GPS points with the three percentiles
pct19_whl <- timeJoin(GPSwhale19, pct19, thresh = 3600, interpolate = FALSE)

#Summarize
buoy19pctSummary <- pct19_whl %>%
  group_by(species) %>%
  summarize(pct10_125_mean = mean(pct10_TOL125),
            pct50_125_mean = mean(pct50_TOL125),
            pct90_125_mean = mean(pct90_TOL125),
            pct10_2000_mean = mean(pct10_TOL2000),
            pct50_2000_mean = mean(pct50_TOL2000),
            pct90_2000_mean = mean(pct90_TOL2000))
```

Buoy 20
```{r}
# Create columns with percentiles and BIND
pct50_20_select <- pct50_20 %>% select(UTC, TOL_125, TOL_2000) %>% 
  rename(pct50_TOL125 = TOL_125,
         pct50_TOL2000 = TOL_2000)
pct10_20_select <- pct10_20 %>% select(UTC, TOL_125, TOL_2000) %>%
   rename(pct10_TOL125 = TOL_125,
         pct10_TOL2000 = TOL_2000)
pct90_20_select <- pct90_20 %>% select(UTC, TOL_125, TOL_2000) %>%
   rename(pct90_TOL125 = TOL_125,
         pct90_TOL2000 = TOL_2000)

# Bind
pct20 <- cbind(pct10_20_select, pct50_20_select, pct90_20_select)

# Join GPS points with the three percentiles
pct20_whl <- timeJoin(GPSwhale20, pct20, thresh = 3600, interpolate = FALSE)

#Summarize
buoy20pctSummary <- pct20_whl %>%
  group_by(species) %>%
  summarize(pct10_125_mean = mean(pct10_TOL125),
            pct50_125_mean = mean(pct50_TOL125),
            pct90_125_mean = mean(pct90_TOL125),
            pct10_2000_mean = mean(pct10_TOL2000),
            pct50_2000_mean = mean(pct50_TOL2000),
            pct90_2000_mean = mean(pct90_TOL2000))
```

Buoy 21
```{r}
# Create columns with percentiles and BIND
pct50_21_select <- pct50_21 %>% select(UTC, TOL_125, TOL_2000) %>% 
  rename(pct50_TOL125 = TOL_125,
         pct50_TOL2000 = TOL_2000)
pct10_21_select <- pct10_21 %>% select(UTC, TOL_125, TOL_2000) %>%
   rename(pct10_TOL125 = TOL_125,
         pct10_TOL2000 = TOL_2000)
pct90_21_select <- pct90_21 %>% select(UTC, TOL_125, TOL_2000) %>%
   rename(pct90_TOL125 = TOL_125,
         pct90_TOL2000 = TOL_2000)

# Bind
pct21 <- cbind(pct10_21_select, pct50_21_select, pct90_21_select)

# Join GPS points with the three percentiles
pct21_whl <- timeJoin(GPSwhale21, pct21, thresh = 3600, interpolate = FALSE)

#Summarize
buoy21pctSummary <- pct21_whl %>%
  group_by(species) %>%
  summarize(pct10_125_mean = mean(pct10_TOL125),
            pct50_125_mean = mean(pct50_TOL125),
            pct90_125_mean = mean(pct90_TOL125),
            pct10_2000_mean = mean(pct10_TOL2000),
            pct50_2000_mean = mean(pct50_TOL2000),
            pct90_2000_mean = mean(pct90_TOL2000))
```

Buoy 22
```{r}
# Create columns with percentiles and BIND
pct50_22_select <- pct50_22 %>% select(UTC, TOL_125, TOL_2000) %>% 
  rename(pct50_TOL125 = TOL_125,
         pct50_TOL2000 = TOL_2000)
pct10_22_select <- pct10_22 %>% select(UTC, TOL_125, TOL_2000) %>%
   rename(pct10_TOL125 = TOL_125,
         pct10_TOL2000 = TOL_2000)
pct90_22_select <- pct90_22 %>% select(UTC, TOL_125, TOL_2000) %>%
   rename(pct90_TOL125 = TOL_125,
         pct90_TOL2000 = TOL_2000)

# Bind
pct22 <- cbind(pct10_22_select, pct50_22_select, pct90_22_select)

# Join GPS points with the three percentiles
pct22_whl <- timeJoin(GPSwhale22, pct22, thresh = 3600, interpolate = FALSE)

#Summarize
buoy22pctSummary <- pct22_whl %>%
  group_by(species) %>%
  summarize(pct10_125_mean = mean(pct10_TOL125),
            pct50_125_mean = mean(pct50_TOL125),
            pct90_125_mean = mean(pct90_TOL125),
            pct10_2000_mean = mean(pct10_TOL2000),
            pct50_2000_mean = mean(pct50_TOL2000),
            pct90_2000_mean = mean(pct90_TOL2000))
```

Buoy 23
```{r}
# Create columns with percentiles and BIND
pct50_23_select <- pct50_23 %>% select(UTC, TOL_125, TOL_2000) %>% 
  rename(pct50_TOL125 = TOL_125,
         pct50_TOL2000 = TOL_2000)
pct10_23_select <- pct10_23 %>% select(UTC, TOL_125, TOL_2000) %>%
   rename(pct10_TOL125 = TOL_125,
         pct10_TOL2000 = TOL_2000)
pct90_23_select <- pct90_23 %>% select(UTC, TOL_125, TOL_2000) %>%
   rename(pct90_TOL125 = TOL_125,
         pct90_TOL2000 = TOL_2000)

# Bind
pct23 <- cbind(pct10_23_select, pct50_23_select, pct90_23_select)

# Join GPS points with the three percentiles
pct23_whl <- timeJoin(GPSwhale23, pct23, thresh = 3600, interpolate = FALSE)

#Summarize
buoy23pctSummary <- pct23_whl %>%
  group_by(species) %>%
  summarize(pct10_125_mean = mean(pct10_TOL125),
            pct50_125_mean = mean(pct50_TOL125),
            pct90_125_mean = mean(pct90_TOL125),
            pct10_2000_mean = mean(pct10_TOL2000),
            pct50_2000_mean = mean(pct50_TOL2000),
            pct90_2000_mean = mean(pct90_TOL2000))
```

Knit all summary percentiles into a table
```{r}

## Buoy 7 ##
sumPct7 <- kable(buoy7pctSummary, col.names = c("Animal", "10th", "Mean", "90th", "10th", "Mean", "90th"),
                                     caption = "Summary Percentiles for Buoy 7") %>%
  kableExtra::kable_classic() %>%
  kableExtra::add_header_above(c(" " = 1, "TOL 125" = 3, "TOL 2000" = 3))

## Buoy 8 ##
sumPct8 <- kable(buoy8pctSummary, col.names = c("Animal", "10th", "Mean", "90th", "10th", "Mean", "90th"),
                                     caption = "Summary Percentiles for Buoy 8") %>%
  kableExtra::kable_classic() %>%
  kableExtra::add_header_above(c(" " = 1, "TOL 125" = 3, "TOL 2000" = 3))

## Buoy 10 ##
sumPct10 <- kable(buoy10pctSummary, col.names = c("Animal", "10th", "Mean", "90th", "10th", "Mean", "90th"),
                                     caption = "Summary Percentiles for Buoy 10") %>%
  kableExtra::kable_classic() %>%
  kableExtra::add_header_above(c(" " = 1, "TOL 125" = 3, "TOL 2000" = 3))

## Buoy 12 ##
sumPct12 <- kable(buoy12pctSummary, col.names = c("Animal", "10th", "Mean", "90th", "10th", "Mean", "90th"),
                                     caption = "Summary Percentiles for Buoy 12") %>%
  kableExtra::kable_classic() %>%
  kableExtra::add_header_above(c(" " = 1, "TOL 125" = 3, "TOL 2000" = 3))

## Buoy 13 ##
sumPct13 <- kable(buoy13pctSummary, col.names = c("Animal", "10th", "Mean", "90th", "10th", "Mean", "90th"),
                                     caption = "Summary Percentiles for Buoy 13") %>%
  kableExtra::kable_classic() %>%
  kableExtra::add_header_above(c(" " = 1, "TOL 125" = 3, "TOL 2000" = 3))

## Buoy 14 ##
#kable(buoy14pctSummary, col.names = c("Animal", "10th", "Mean", "90th", "10th", "Mean", "90th"),
#                                     caption = "Summary Percentiles for Buoy 14") %>%
#  kableExtra::kable_classic() %>%
#  kableExtra::add_header_above(c(" " = 1, "TOL 125" = 3, "TOL 2000" = 3))

## Buoy 16 ##
sumPct16 <- kable(buoy16pctSummary, col.names = c("Animal", "10th", "Mean", "90th", "10th", "Mean", "90th"),
                                     caption = "Summary Percentiles for Buoy 16") %>%
  kableExtra::kable_classic() %>%
  kableExtra::add_header_above(c(" " = 1, "TOL 125" = 3, "TOL 2000" = 3))

## Buoy 18 ##
sumPct18 <- kable(buoy18pctSummary, col.names = c("Animal", "10th", "Mean", "90th", "10th", "Mean", "90th"),
                                     caption = "Summary Percentiles for Buoy 18") %>%
  kableExtra::kable_classic() %>%
  kableExtra::add_header_above(c(" " = 1, "TOL 125" = 3, "TOL 2000" = 3))

## Buoy 19 ##
sumPct19 <- kable(buoy19pctSummary, col.names = c("Animal", "10th", "Mean", "90th", "10th", "Mean", "90th"),
                                     caption = "Summary Percentiles for Buoy 19") %>%
  kableExtra::kable_classic() %>%
  kableExtra::add_header_above(c(" " = 1, "TOL 125" = 3, "TOL 2000" = 3))

## Buoy 20 ##
sumPct20 <- kable(buoy20pctSummary, col.names = c("Animal", "10th", "Mean", "90th", "10th", "Mean", "90th"),
                                     caption = "Summary Percentiles for Buoy 20") %>%
  kableExtra::kable_classic() %>%
  kableExtra::add_header_above(c(" " = 1, "TOL 125" = 3, "TOL 2000" = 3))

## Buoy 21 ##
sumPct21 <- kable(buoy21pctSummary, col.names = c("Animal", "10th", "Mean", "90th", "10th", "Mean", "90th"),
                                     caption = "Summary Percentiles for Buoy 21") %>%
  kableExtra::kable_classic() %>%
  kableExtra::add_header_above(c(" " = 1, "TOL 125" = 3, "TOL 2000" = 3))

## Buoy 22 ##
sumPct22 <- kable(buoy22pctSummary, col.names = c("Animal", "10th", "Mean", "90th", "10th", "Mean", "90th"),
                                     caption = "Summary Percentiles for Buoy 22") %>%
  kableExtra::kable_classic() %>%
  kableExtra::add_header_above(c(" " = 1, "TOL 125" = 3, "TOL 2000" = 3))

## Buoy 23 ##
sumPct23 <- kable(buoy23pctSummary, col.names = c("Animal", "10th", "Mean", "90th", "10th", "Mean", "90th"),
                                     caption = "Summary Percentiles for Buoy 23") %>%
  kableExtra::kable_classic() %>%
  kableExtra::add_header_above(c(" " = 1, "TOL 125" = 3, "TOL 2000" = 3))

```

Time Series Heat Map
```{r fig.width=12}

pdf(paste("timeSeries8-31",".pdf"), width=14, height=10)

### Buoy 7 ####
W.07 <- fulltrack07 %>%
  select(UTC, BWpresence, species) %>%
  filter(species == "ZC")

W.07plot <- plotPresGrid(W.07, bin = 'hour', type = 'presence')

g1 <- (plotTimeseries(fulltrack07, column='TOL_125', bin='1hour', style = 'heatmap', title='Buoy 7: TOL 125 Plot + ZC'))
g2 <- plotPresGrid(W.07, bin = 'hour', type = 'presence', start = min(fulltrack07$UTC),
                         end = max(fulltrack07$UTC), fill = NA, color = "white")

# Create plot with TOL125
combined.g <- g1
combined.g$layers <- c(combined.g$layers, g2$layers)

t1 <- plotTimeseries(fulltrack07, column='TOL_2000', bin='1hour', style = 'heatmap', title='Buoy 7: TOL 2000 Plot + ZC')
t2 <- plotPresGrid(W.07, bin = 'hour', type = 'presence', start = min(fulltrack07$UTC),
                         end = max(fulltrack07$UTC), fill = NA, color = "white")

# Create plot with TOL2000
combined.t <- t1
combined.t$layers <- c(combined.t$layers, t2$layers)

combined.g + combined.t
```

```{r fig.width=12}
###### Buoy 8 ####
W.08 <- fulltrack08 %>%
  select(UTC, BWpresence, species) %>%
  filter(species == "ZC")

W.08plot <- plotPresGrid(W.08, bin = 'hour', type = 'presence')

g1 <- (plotTimeseries(fulltrack08, column='TOL_125', bin='1hour', style = 'heatmap', title='Buoy 8: TOL 125 Plot + ZC'))
g2 <- plotPresGrid(W.08, bin = 'hour', type = 'presence', start = min(fulltrack08$UTC),
                         end = max(fulltrack08$UTC), fill = NA, color = "white")

combined.g <- g1
combined.g$layers <- c(combined.g$layers, g2$layers)
#####
t1 <- plotTimeseries(fulltrack08, column='TOL_2000', bin='1hour', style = 'heatmap', title='Buoy 8: TOL 2000 Plot + ZC')
t2 <- plotPresGrid(W.08, bin = 'hour', type = 'presence', start = min(fulltrack08$UTC),
                         end = max(fulltrack08$UTC), fill = NA, color = "white")

combined.t <- t1
combined.t$layers <- c(combined.t$layers, t2$layers)

combined.g + combined.t
```

```{r fig.width=12}
#### Buoy 10  ######

W.10 <- fulltrack10 %>%
  select(UTC, BWpresence, species) %>%
  filter(species == "ZC")

W.10plot <- plotPresGrid(W.10, bin = 'hour', type = 'presence')

g1 <- (plotTimeseries(fulltrack10, column='TOL_125', bin='1hour', style = 'heatmap', title='Buoy 10: TOL 125 Plot + ZC'))
g2 <- plotPresGrid(W.10, bin = 'hour', type = 'presence', start = min(fulltrack10$UTC),
                         end = max(fulltrack10$UTC), fill = NA, color = "white")

combined.g <- g1
combined.g$layers <- c(combined.g$layers, g2$layers)
#####
t1 <- plotTimeseries(fulltrack10, column='TOL_2000', bin='1hour', style = 'heatmap', title='Buoy 10: TOL 2000 Plot + ZC')
t2 <- plotPresGrid(W.10, bin = 'hour', type = 'presence', start = min(fulltrack10$UTC),
                         end = max(fulltrack10$UTC), fill = NA, color = "white")

combined.t <- t1
combined.t$layers <- c(combined.t$layers, t2$layers)

combined.g + combined.t
```

```{r fig.width=12}
## Buoy 12
W.12 <- fulltrack12 %>%
  select(UTC, BWpresence, species) %>%
  filter(species == "ZC")

W.12plot <- plotPresGrid(W.12, bin = 'hour', type = 'presence')

g1 <- (plotTimeseries(fulltrack12, column='TOL_125', bin='1hour', style = 'heatmap', title='Buoy 12: TOL 125 Plot + ZC'))
g2 <- plotPresGrid(W.12, bin = 'hour', type = 'presence', start = min(fulltrack12$UTC),
                         end = max(fulltrack12$UTC), fill = NA, color = "white")

combined.g <- g1
combined.g$layers <- c(combined.g$layers, g2$layers)
#####
t1 <- plotTimeseries(fulltrack12, column='TOL_2000', bin='1hour', style = 'heatmap', title='Buoy 12: TOL 2000 Plot + ZC')
t2 <- plotPresGrid(W.12, bin = 'hour', type = 'presence', start = min(fulltrack12$UTC),
                         end = max(fulltrack12$UTC), fill = NA, color = "white")

combined.t <- t1
combined.t$layers <- c(combined.t$layers, t2$layers)

combined.g + combined.t
```

```{r fig.width=12}
## Buoy 13
W.13 <- fulltrack13 %>%
  select(UTC, BWpresence, species) %>%
  filter(species == "ZC")

W.13plot <- plotPresGrid(W.13, bin = 'hour', type = 'presence')

g1 <- (plotTimeseries(fulltrack13, column='TOL_125', bin='1hour', style = 'heatmap', title='Buoy 13: TOL 125 Plot + ZC'))
g2 <- plotPresGrid(W.13, bin = 'hour', type = 'presence', start = min(fulltrack13$UTC),
                         end = max(fulltrack13$UTC), fill = NA, color = "white")

combined.g <- g1
combined.g$layers <- c(combined.g$layers, g2$layers)
#####
t1 <- plotTimeseries(fulltrack13, column='TOL_2000', bin='1hour', style = 'heatmap', title='Buoy 13: TOL 2000 Plot + ZC')
t2 <- plotPresGrid(W.13, bin = 'hour', type = 'presence', start = min(fulltrack13$UTC),
                         end = max(fulltrack13$UTC), fill = NA, color = "white")

combined.t <- t1
combined.t$layers <- c(combined.t$layers, t2$layers)

combined.g + combined.t
```

```{r fig.width=12}
## Buoy 14
W.14 <- fulltrack14 %>%
  select(UTC, BWpresence, species) %>%
  filter(species == "ZC")

W.14plot <- plotPresGrid(W.14, bin = 'hour', type = 'presence')

g1 <- (plotTimeseries(fulltrack14, column='TOL_125', bin='hour', style = 'heatmap', title='Buoy 14: TOL 125 Plot + ZC'))
g2 <- plotPresGrid(W.14, bin = 'hour', type = 'presence', start = min(fulltrack14$UTC),
                         end = max(fulltrack14$UTC), fill = NA, color = "white")

combined.g <- g1
combined.g$layers <- c(combined.g$layers, g2$layers)
#####
t1 <- plotTimeseries(fulltrack14, column='TOL_2000', bin='hour', style = 'heatmap', title='Buoy 14: TOL 2000 Plot + ZC')
t2 <- plotPresGrid(W.14, bin = 'hour', type = 'presence', start = min(fulltrack14$UTC),
                         end = max(fulltrack14$UTC), fill = NA, color = "white")

combined.t <- t1
combined.t$layers <- c(combined.t$layers, t2$layers)

combined.g + combined.t
```

```{r fig.width=12}
## Buoy 16
W.16 <- fulltrack16 %>%
  select(UTC, BWpresence, species) %>%
  filter(species == "ZC")

W.16plot <- plotPresGrid(W.16, bin = 'hour', type = 'presence')

g1 <- (plotTimeseries(fulltrack16, column='TOL_125', bin='1hour', style = 'heatmap', title='Buoy 16: TOL 125 Plot + ZC'))
g2 <- plotPresGrid(W.16, bin = 'hour', type = 'presence', start = min(fulltrack16$UTC),
                         end = max(fulltrack16$UTC), fill = NA, color = "white")

combined.g <- g1
combined.g$layers <- c(combined.g$layers, g2$layers)
#####
t1 <- plotTimeseries(fulltrack16, column='TOL_2000', bin='1hour', style = 'heatmap', title='Buoy 16: TOL 2000 Plot + ZC')
t2 <- plotPresGrid(W.16, bin = 'hour', type = 'presence', start = min(fulltrack16$UTC),
                         end = max(fulltrack16$UTC), fill = NA, color = "white")

combined.t <- t1
combined.t$layers <- c(combined.t$layers, t2$layers)

combined.g + combined.t
```

```{r fig.width=12}
## Buoy 18
W.18 <- fulltrack18 %>%
  select(UTC, BWpresence, species) %>%
  filter(species == "ZC")

W.18plot <- plotPresGrid(W.18, bin = 'hour', type = 'presence')

g1 <- (plotTimeseries(fulltrack18, column='TOL_125', bin='1hour', style = 'heatmap', title='Buoy 18: TOL 125 Plot + ZC'))
g2 <- plotPresGrid(W.18, bin = 'hour', type = 'presence', start = min(fulltrack18$UTC),
                         end = max(fulltrack18$UTC), fill = NA, color = "white")

combined.g <- g1
combined.g$layers <- c(combined.g$layers, g2$layers)
#####
t1 <- plotTimeseries(fulltrack18, column='TOL_2000', bin='1hour', style = 'heatmap', title='Buoy 18: TOL 2000 Plot + ZC')
t2 <- plotPresGrid(W.18, bin = 'hour', type = 'presence', start = min(fulltrack18$UTC),
                         end = max(fulltrack18$UTC), fill = NA, color = "white")

combined.t <- t1
combined.t$layers <- c(combined.t$layers, t2$layers)

combined.g + combined.t
```

```{r fig.width=12}
## Buoy 19
W.19 <- fulltrack19 %>%
  select(UTC, BWpresence, species) %>%
  filter(species == "ZC")

W.19plot <- plotPresGrid(W.19, bin = 'hour', type = 'presence')

g1 <- (plotTimeseries(fulltrack19, column='TOL_125', bin='1hour', style = 'heatmap', title='Buoy 19: TOL 125 Plot + ZC'))
g2 <- plotPresGrid(W.19, bin = 'hour', type = 'presence', start = min(fulltrack19$UTC),
                         end = max(fulltrack19$UTC), fill = NA, color = "white")

combined.g <- g1
combined.g$layers <- c(combined.g$layers, g2$layers)
#####
t1 <- plotTimeseries(fulltrack19, column='TOL_2000', bin='1hour', style = 'heatmap', title='Buoy 19: TOL 2000 Plot + ZC')
t2 <- plotPresGrid(W.19, bin = 'hour', type = 'presence', start = min(fulltrack19$UTC),
                         end = max(fulltrack19$UTC), fill = NA, color = "white")

combined.t <- t1
combined.t$layers <- c(combined.t$layers, t2$layers)

combined.g + combined.t
```

```{r fig.width=12}
## Buoy 20
W.20 <- fulltrack20 %>%
  select(UTC, BWpresence, species) %>%
  filter(species == "ZC")

W.20plot <- plotPresGrid(W.20, bin = 'hour', type = 'presence')

g1 <- (plotTimeseries(fulltrack20, column='TOL_125', bin='1hour', style = 'heatmap', title='Buoy 20: TOL 125 Plot + ZC'))
g2 <- plotPresGrid(W.20, bin = 'hour', type = 'presence', start = min(fulltrack20$UTC),
                         end = max(fulltrack20$UTC), fill = NA, color = "white")

combined.g <- g1
combined.g$layers <- c(combined.g$layers, g2$layers)
#####
t1 <- plotTimeseries(fulltrack20, column='TOL_2000', bin='1hour', style = 'heatmap', title='Buoy 20: TOL 2000 Plot + ZC')
t2 <- plotPresGrid(W.20, bin = 'hour', type = 'presence', start = min(fulltrack20$UTC),
                         end = max(fulltrack20$UTC), fill = NA, color = "white")

combined.t <- t1
combined.t$layers <- c(combined.t$layers, t2$layers)

combined.g + combined.t
```

```{r fig.width=12}
## Buoy 21
W.21 <- fulltrack21 %>%
  select(UTC, BWpresence, species) %>%
  filter(species == "ZC")

W.21plot <- plotPresGrid(W.21, bin = 'hour', type = 'presence')

g1 <- (plotTimeseries(fulltrack21, column='TOL_125', bin='1hour', style = 'heatmap', title='Buoy 21: TOL 125 Plot + ZC'))
g2 <- plotPresGrid(W.21, bin = 'hour', type = 'presence', start = min(fulltrack21$UTC),
                         end = max(fulltrack21$UTC), fill = NA, color = "white")

combined.g <- g1
combined.g$layers <- c(combined.g$layers, g2$layers)
#####
t1 <- plotTimeseries(fulltrack21, column='TOL_2000', bin='1hour', style = 'heatmap', title='Buoy 21: TOL 2000 Plot + ZC')
t2 <- plotPresGrid(W.21, bin = 'hour', type = 'presence', start = min(fulltrack21$UTC),
                         end = max(fulltrack21$UTC), fill = NA, color = "white")

combined.t <- t1
combined.t$layers <- c(combined.t$layers, t2$layers)

combined.g + combined.t
```

```{r fig.width=12}
## Buoy 12
W.22 <- fulltrack22 %>%
  select(UTC, BWpresence, species) %>%
  filter(species == "ZC")

W.22plot <- plotPresGrid(W.22, bin = 'hour', type = 'presence')

g1 <- (plotTimeseries(fulltrack22, column='TOL_125', bin='1hour', style = 'heatmap', title='Buoy 22: TOL 125 Plot + ZC'))
g2 <- plotPresGrid(W.22, bin = 'hour', type = 'presence', start = min(fulltrack22$UTC),
                         end = max(fulltrack22$UTC), fill = NA, color = "white")

combined.g <- g1
combined.g$layers <- c(combined.g$layers, g2$layers)
#####
t1 <- plotTimeseries(fulltrack22, column='TOL_2000', bin='1hour', style = 'heatmap', title='Buoy 22: TOL 2000 Plot + ZC')
t2 <- plotPresGrid(W.22, bin = 'hour', type = 'presence', start = min(fulltrack22$UTC),
                         end = max(fulltrack22$UTC), fill = NA, color = "white")

combined.t <- t1
combined.t$layers <- c(combined.t$layers, t2$layers)

combined.g + combined.t
```

```{r fig.width=12}
## Buoy 23
W.23 <- fulltrack23 %>%
  select(UTC, BWpresence, species) %>%
  filter(species == "ZC")

W.23plot <- plotPresGrid(W.23, bin = 'hour', type = 'presence')

g1 <- (plotTimeseries(fulltrack23, column='TOL_125', bin='1hour', style = 'heatmap', title='Buoy 23: TOL 125 Plot + ZC'))
g2 <- plotPresGrid(W.23, bin = 'hour', type = 'presence', start = min(fulltrack23$UTC),
                         end = max(fulltrack23$UTC), fill = NA, color = "white")

combined.g <- g1
combined.g$layers <- c(combined.g$layers, g2$layers)
#####
t1 <- plotTimeseries(fulltrack23, column='TOL_2000', bin='1hour', style = 'heatmap', title='Buoy 23: TOL 2000 Plot + ZC')
t2 <- plotPresGrid(W.23, bin = 'hour', type = 'presence', start = min(fulltrack23$UTC),
                         end = max(fulltrack23$UTC), fill = NA, color = "white")

combined.t <- t1
combined.t$layers <- c(combined.t$layers, t2$layers)

combined.g + combined.t
```

```{r}
save.image(file = ".RData")
```

plot whale presence and SS side by side. Stacked?
```{r}
W.08 <- fulltrack08 %>%
  select(UTC, BWpresence, species) %>%
  filter(species == "ZC")

W.08plot <- plotPresGrid(W.08, bin = 'hour', type = 'presence')

plotTimeseries(fulltrack08, column='TOL_2000', bin='1hour', style = 'heatmap', title='Buoy 8: TOL 2000 Plot') +  W.08plot

#############

W.10 <- fulltrack10 %>%
  select(UTC, BWpresence, species) %>%
  filter(species == "ZC")

W.10plot <- plotPresGrid(W.10, bin = 'hour', type = 'presence', start = min(fulltrack10$UTC),
                         end = max(fulltrack10$UTC), by = 'BWpresence', alpha = 0.01, color = "red")

plotTimeseries(fulltrack10, column='TOL_2000', bin='1hour', style = 'heatmap', title='Buoy 10: TOL 2000 Plot') +  W.10plot
plotTimeseries(fulltrack10, column='TOL_125', bin='1hour', style = 'heatmap', title='Buoy 10: TOL 125 Plot') +  W.10plot
#############

W.13 <- fulltrack13 %>%
  select(UTC, BWpresence, species) %>%
  filter(species == "ZC")

W.13plot <- plotPresGrid(W.13, bin = 'hour', type = 'presence', start = min(fulltrack13$UTC),
                         end = max(fulltrack13$UTC), by = 'BWpresence', alpha = 0.5)

plotTimeseries(fulltrack13, column='TOL_2000', bin='1hour', style = 'heatmap', title='Buoy 13: TOL 2000 Plot') +  W.13plot
plotTimeseries(fulltrack13, column='TOL_125', bin='1hour', style = 'heatmap', title='Buoy 13: TOL 125 Plot') +  W.13plot
##############

W.20 <- fulltrack20 %>%
  select(UTC, BWpresence, species) %>%
  filter(species == "ZC")

W.20plot <- plotPresGrid(W.20, bin = 'hour', type = 'presence', start = min(fulltrack20$UTC),
                         end = max(fulltrack20$UTC),  by = 'BWpresence', alpha = 0.5)

plotTimeseries(fulltrack20, column='TOL_2000', bin='1hour', style = 'heatmap', title='Buoy 20: TOL 2000 Plot') +  W.20plot
plotTimeseries(fulltrack20, column='TOL_125', bin='1hour', style = 'heatmap', title='Buoy 20: TOL 125 Plot') +  W.20plot

##########

g1 <- plotTimeseries(fulltrack10, column='TOL_2000', bin='1hour', style = 'heatmap', title='Buoy 10: TOL 2000 Plot')
g2 <- plotPresGrid(W.10, bin = 'hour', type = 'presence', start = min(fulltrack10$UTC),
                         end = max(fulltrack10$UTC), fill = NA, color = "white")
S.10 <- fulltrack10 %>% select(UTC, shipin10km) %>% filter(shipin10km == "1")
g3 <- plotPresGrid(S.10, bin = 'hour', type = 'presence', start = min(fulltrack10$UTC),
                         end = max(fulltrack10$UTC), fill = NA, color = "red")

combined <- g1
combined$layers <- c(combined$layers, g2$layers, g3$layers)
combined
```

Bind ALL DRIFTS into one dataframe
```{r}
allDrifts <- rbind(fulltrack07, fulltrack08, fulltrack10, 
                   fulltrack12, fulltrack13, fulltrack14, 
                   fulltrack16, fulltrack18, fulltrack19, 
                   fulltrack20, fulltrack21, fulltrack22, fulltrack23)
```

Where are loudest sound levels? See @haver2020 for details on percentiles
5th percentile = exceeded most of the time, so refers to background sound
90th percentile = exceeded 10% of the time, so refers to PEAKS in sound level
Also see @haver2021 (fig 2) for daily TOL levels by site (can do by buoy)
```{r}

pctlist <- c("07", "08", "10", "12", "14", "16", "18", "19",
             "20", "21", "22", "23")
# Doesn't show time stamp!
for (pc in pctlist) {
  data <- get(paste0("pct99_", pc)) %>%
    summarise_if(is.numeric, list(~ max(., na.rm=TRUE)))

  assign(paste0("max_", pc), data)
}

print(max_07)

## Add multi. Sum up rows and extract min/max
ggplot(data = TO_pct01_07, aes()) + 
  geom_histogram()

summary(TO_pct01_07)
```

Where are the quietest sound levels by band?
```{r}
pctlist <- c("07", "08", "10", "12", "14", "16", "18", "19",
             "20", "21", "22", "23")
# Doesn't show time stamp!
for (pc in pctlist) {
  data <- get(paste0("pct01_", pc)) %>%
    summarise_if(is.numeric, list(~ min(., na.rm=TRUE)))

  assign(paste0("min_", pc), data)
}

print(min_07)
```

Where do we hear the most whale detections? Which buoy had the most detections?
```{r}
# Summarize number of detections by species
allDrifts %>%
  group_by(species) %>%
  summarize(n())
## ZC had the most

#Which buoy had the most detections?
allDrifts %>%
  count(station, BWpresence) %>%
  filter(!BWpresence == 0) %>%
  arrange(desc(n))
## Buoy 18 has the most

#Plot buoys
ggplot() +
     geom_path(data=allDrifts, aes(x=Longitude, y=Latitude, group=factor(station), col=factor(station))) 
```

Summary of each species: highest/lowest sound levels, avg sound levels they're exposed to
```{r}
# Sort sperm whale detection by decreasing BB sound levels
allDrifts %>%
  dplyr::filter(species == "PM") %>%
  arrange(desc(BB_20.24000))

PM <- allDrifts %>%
  dplyr::filter(species == "PM")

PM %>%
  group_by(station) %>%
  summarize(mean = mean(TOL_2000), n = n())

print(paste("PM Max BB sound level: ", max(PM$BB_20.24000, na.rm = TRUE)))
print(paste("PM Min BB sound level: ", min(PM$BB_20.24000, na.rm = TRUE)))
print(paste("PM Max TOL125 sound level: ", max(PM$TOL_125, na.rm = TRUE)))
print(paste("PM Min TOL125 sound level: ", min(PM$TOL_125, na.rm = TRUE)))
print(paste("PM Max TOL20000 sound level: ", max(PM$TOL_20000, na.rm = TRUE)))
print(paste("PM Min TOL20000 sound level: ", min(PM$TOL_20000, na.rm = TRUE)))

# One detection value for BWC
BWC <- allDrifts %>%
  dplyr::filter(species == "BWC")
print(paste("BWC Max BB sound level: ", max(BWC$BB_20.24000, na.rm = TRUE)))
print(paste("BWC Min BB sound level: ", min(BWC$BB_20.24000, na.rm = TRUE)))
print(paste("BWC Max TOL125 sound level: ", max(BWC$TOL_125, na.rm = TRUE)))
print(paste("BWC Min TOL125 sound level: ", min(BWC$TOL_125, na.rm = TRUE)))
print(paste("BWC Max TOL20000 sound level: ", max(BWC$TOL_20000, na.rm = TRUE)))
print(paste("BWC Min TOL20000 sound level: ", min(BWC$TOL_20000, na.rm = TRUE)))

BB <- allDrifts %>%
  dplyr::filter(species == "BB")
print(paste("BB Max BB sound level: ", max(BB$BB_20.24000, na.rm = TRUE)))
print(paste("BB Min BB sound level: ", min(BB$BB_20.24000, na.rm = TRUE)))
print(paste("BB Max TOL125 sound level: ", max(BB$TOL_125, na.rm = TRUE)))
print(paste("BB Min TOL125 sound level: ", min(BB$TOL_125, na.rm = TRUE)))
print(paste("BB Max TOL20000 sound level: ", max(BB$TOL_20000, na.rm = TRUE)))
print(paste("BB Min TOL20000 sound level: ", min(BB$TOL_20000, na.rm = TRUE)))

BW37V <- allDrifts %>%
  dplyr::filter(species == "BW37V")
print(paste("BW37V Max BB sound level: ", max(BW37V$BB_20.24000, na.rm = TRUE)))
print(paste("BW37V Min BB sound level: ", min(BW37V$BB_20.24000, na.rm = TRUE)))
print(paste("BW37V Max TOL125 sound level: ", max(BW37V$TOL_125, na.rm = TRUE)))
print(paste("BW37V Min TOL125 sound level: ", min(BW37V$TOL_125, na.rm = TRUE)))
print(paste("BW37V Max TOL20000 sound level: ", max(BW37V$TOL_20000, na.rm = TRUE)))
print(paste("BW37V Min TOL20000 sound level: ", min(BW37V$TOL_20000, na.rm = TRUE)))

BW43 <- allDrifts %>%
  dplyr::filter(species == "BW43")
print(paste("BW43 Max BB sound level: ", max(BW43$BB_20.24000, na.rm = TRUE)))
print(paste("BW43 Min BB sound level: ", min(BW43$BB_20.24000, na.rm = TRUE)))
print(paste("BW43 Max TOL125 sound level: ", max(BW43$TOL_125, na.rm = TRUE)))
print(paste("BW43 Min TOL125 sound level: ", min(BW43$TOL_125, na.rm = TRUE)))
print(paste("BW43 Max TOL20000 sound level: ", max(BW43$TOL_20000, na.rm = TRUE)))
print(paste("BW43 Min TOL20000 sound level: ", min(BW43$TOL_20000, na.rm = TRUE)))

ZC <- allDrifts %>%
  dplyr::filter(species == "ZC")
print(paste("ZC Max BB sound level: ", max(ZC$BB_20.24000, na.rm = TRUE)))
print(paste("ZC Min BB sound level: ", min(ZC$BB_20.24000, na.rm = TRUE)))
print(paste("ZC Max TOL125 sound level: ", max(ZC$TOL_125, na.rm = TRUE)))
print(paste("ZC Min TOL125 sound level: ", min(ZC$TOL_125, na.rm = TRUE)))
print(paste("ZC Max TOL20000 sound level: ", max(ZC$TOL_20000, na.rm = TRUE)))
print(paste("ZC Min TOL20000 sound level: ", min(ZC$TOL_20000, na.rm = TRUE)))

```
Number of detections
```{r}
print(paste("Number of PM detections: ", nrow(PM)))
print(paste("Number of ZC detections: ", nrow(ZC)))
print(paste("Number of BW43 detections: ", nrow(BW43)))
print(paste("Number of BWC detections: ", nrow(BWC)))
print(paste("Number of BB detections: ", nrow(BB)))
print(paste("Number of BW37V detections: ", nrow(BW37V)))

#Plot
sp <- allDrifts %>% filter(!species == 'NA')
ggplot(sp, aes(species)) +
  geom_bar()
```
How do I find WHERE the highest/lowest detections are happening?
```{r}
ZC <- allDrifts %>%
  dplyr::filter(species == "ZC")

ZC[which.max(ZC$BB_20.24000),]
ZCmax2000 <- ZC[which.max(ZC$TOL_2000),]

#Point of ZC max TOL2000 sound level. Need to put in context of map.
ggplot(data = ZCmax2000, aes(x = Longitude, y = Latitude)) +
  geom_point()

#not working
allDrifts %>%
  group_by(species) %>%
  filter(max(allDrifts$BB_20.24000))

allDrifts %>%
  summarise_if(is.numeric, list(~ min(., na.rm=TRUE)))

```