### 2023-09-13 02_AIS.Rmd
Update 02_AIS.Rmd
reran with updated GPS pts

---
title: "AIS"
author: "Marina Bozinovic"
date: "2023-04-11"
output: html_document
---

#Introduction
AIS download from Marine Cadastre and match to buoy datasets.

Required Packages
```{r, }
library(PAMscapes)
library(tidyverse)
```

-----------------------
From Taiki's AIS Workflow
```{r}
#marcadFiles07 <- downloadMarCadAIS(GPSwhale7, outDir = "D:/AIS", overwrite = FALSE, unzip = TRUE)
#marcadFiles08 <- downloadMarCadAIS(GPSwhale8, outDir = "D:/AIS", overwrite = FALSE, unzip = TRUE)
#marcadFiles10 <- downloadMarCadAIS(GPSwhale10, outDir = "D:/AIS", overwrite = FALSE, unzip = TRUE)
#marcadFiles12 <- downloadMarCadAIS(GPSwhale12, outDir = "D:/AIS", overwrite = FALSE, unzip = TRUE)
#marcadFiles13 <- downloadMarCadAIS(GPSwhale13, outDir = "D:/AIS", overwrite = FALSE, unzip = TRUE)
#marcadFiles14 <- downloadMarCadAIS(GPSwhale14, outDir = "D:/AIS", overwrite = FALSE, unzip = TRUE)
#marcadFiles16 <- downloadMarCadAIS(GPSwhale16, outDir = "D:/AIS", overwrite = FALSE, unzip = TRUE)
#marcadFiles18 <- downloadMarCadAIS(GPSwhale18, outDir = "D:/AIS", overwrite = FALSE, unzip = TRUE)
#marcadFiles19 <- downloadMarCadAIS(GPSwhale19, outDir = "D:/AIS", overwrite = FALSE, unzip = TRUE)
#marcadFiles20 <- downloadMarCadAIS(GPSwhale20, outDir = "D:/AIS", overwrite = FALSE, unzip = TRUE)
#marcadFiles21 <- downloadMarCadAIS(GPSwhale21, outDir = "D:/AIS", overwrite = FALSE, unzip = TRUE)
#marcadFiles22 <- downloadMarCadAIS(GPSwhale22, outDir = "D:/AIS", overwrite = FALSE, unzip = TRUE)
#marcadFiles23 <- downloadMarCadAIS(GPSwhale23, outDir = "D:/AIS", overwrite = FALSE, unzip = TRUE)
```

Subset to US West Coast
```{r}
subsetMarCadAIS(inDir="D:/AIS", outDir="D:/AIS/AIS_West", name='West_', 
                latRange=c(20,50), lonRange=c(-140,-110),
                overwrite=FALSE, progress=TRUE)
```
Now read into R only ais points that come in contact with buoys
```{r}
#ais_near_07 <- readLocalAIS(GPSwhale7, aisDir="D:/AIS/AIS_West", distance=10e3) %>%
#   filter(vesselLength >= 100)  #remove vessels under 100m in length

#ais_near_08 <- readLocalAIS(GPSwhale8, aisDir="D:/AIS/AIS_West", distance=10e3) %>%
#   filter(vesselLength >= 100)  #remove vessels under 100m in length

#ais_near_10 <- readLocalAIS(GPSwhale10, aisDir="D:/AIS/AIS_West", distance=10e3) %>%
#   filter(vesselLength >= 100)  #remove vessels under 100m in length

#ais_near_12 <- readLocalAIS(GPSwhale12, aisDir="D:/AIS/AIS_West", distance=10e3) %>%
#   filter(vesselLength >= 100)  #remove vessels under 100m in length

#ais_near_13 <- readLocalAIS(GPSwhale13, aisDir="D:/AIS/AIS_West", distance=10e3) %>%
#   filter(vesselLength >= 100)  #remove vessels under 100m in length

#ais_near_14 <- readLocalAIS(GPSwhale14, aisDir="D:/AIS/AIS_West", distance=10e3) %>%
#   filter(vesselLength >= 100)  #remove vessels under 100m in length

#ais_near_16 <- readLocalAIS(GPSwhale16, aisDir="D:/AIS/AIS_West", distance=10e3) %>%
#   filter(vesselLength >= 100)  #remove vessels under 100m in length

#ais_near_18 <- readLocalAIS(GPSwhale18, aisDir="D:/AIS/AIS_West", distance=10e3) %>%
#   filter(vesselLength >= 100)  #remove vessels under 100m in length

#ais_near_19 <- readLocalAIS(GPSwhale19, aisDir="D:/AIS/AIS_West", distance=10e3) %>%
#   filter(vesselLength >= 100)  #remove vessels under 100m in length

#ais_near_20 <- readLocalAIS(GPSwhale20, aisDir="D:/AIS/AIS_West", distance=10e3) %>%
#   filter(vesselLength >= 100)  #remove vessels under 100m in length

#ais_near_21 <- readLocalAIS(GPSwhale21, aisDir="D:/AIS/AIS_West", distance=10e3) %>%
#   filter(vesselLength >= 100)  #remove vessels under 100m in length
#empty table because no large vessels interacted with this buoy

#ais_near_22 <- readLocalAIS(GPSwhale22, aisDir="D:/AIS/AIS_West", distance=10e3) %>%
#     filter(vesselLength >= 100)  #remove vessels under 100m in length

#ais_near_23 <- readLocalAIS(GPSwhale23, aisDir="D:/AIS/AIS_West", distance=10e3) %>%
#   filter(vesselLength >= 100)  #remove vessels under 100m in length
```

Save `ais_near_` for all buoys
```{r}
#saveRDS(ais_near_07, file = "D:AIS/ais_near_07.rda")
#saveRDS(ais_near_08, file = "D:AIS/ais_near_08.rda")
#saveRDS(ais_near_10, file = "D:AIS/ais_near_10.rda")
#saveRDS(ais_near_12, file = "D:AIS/ais_near_12.rda")
#saveRDS(ais_near_13, file = "D:AIS/ais_near_13.rda")
#saveRDS(ais_near_14, file = "D:AIS/ais_near_14.rda")
#saveRDS(ais_near_16, file = "D:AIS/ais_near_16.rda")
#saveRDS(ais_near_18, file = "D:AIS/ais_near_18.rda")
#saveRDS(ais_near_19, file = "D:AIS/ais_near_19.rda")
#saveRDS(ais_near_20, file = "D:AIS/ais_near_20.rda")
#saveRDS(ais_near_21, file = "D:AIS/ais_near_21.rda")
#saveRDS(ais_near_22, file = "D:AIS/ais_near_22.rda")
#saveRDS(ais_near_23, file = "D:AIS/ais_near_23.rda")
```

Load rds files into R with AIS points near buoys
```{r}
ais_near_07 <- readRDS("D:/AIS/ais_near_07.rda")
ais_near_08 <- readRDS('D:/AIS/ais_near_08.rda')
ais_near_10 <- readRDS('D:/AIS/ais_near_10.rda')
ais_near_12 <- readRDS('D:/AIS/ais_near_12.rda')
ais_near_13 <- readRDS('D:/AIS/ais_near_13.rda')
ais_near_14 <- readRDS('D:/AIS/ais_near_14.rda')
ais_near_16 <- readRDS('D:/AIS/ais_near_16.rda')
ais_near_18 <- readRDS('D:/AIS/ais_near_18.rda')
ais_near_19 <- readRDS('D:/AIS/ais_near_19.rda')
ais_near_20 <- readRDS('D:/AIS/ais_near_20.rda')
ais_near_21 <- readRDS('D:/AIS/ais_near_21.rda')
ais_near_22 <- readRDS('D:/AIS/ais_near_22.rda')
ais_near_23 <- readRDS('D:/AIS/ais_near_23.rda')
```

Match AIS points with buoy points
```{r}
gpsClose07 <- addAIS(GPSwhale7, ais_near_07, interpType = 'close', interpTime = 120)
gpsClose08 <- addAIS(GPSwhale8, ais_near_08, interpType = 'close', interpTime = 120)
gpsClose10 <- addAIS(GPSwhale10, ais_near_10, interpType = 'close', interpTime = 120)
gpsClose12 <- addAIS(GPSwhale12, ais_near_12, interpType = 'close', interpTime = 120)
gpsClose13 <- addAIS(GPSwhale13, ais_near_13, interpType = 'close', interpTime = 120)
gpsClose14 <- addAIS(GPSwhale14, ais_near_14, interpType = 'close', interpTime = 120)
gpsClose16 <- addAIS(GPSwhale16, ais_near_16, interpType = 'close', interpTime = 120)
gpsClose18 <- addAIS(GPSwhale18, ais_near_18, interpType = 'close', interpTime = 120)
gpsClose19 <- addAIS(GPSwhale19, ais_near_19, interpType = 'close', interpTime = 120)
gpsClose20 <- addAIS(GPSwhale20, ais_near_20, interpType = 'close', interpTime = 120)
gpsClose21 <- addAIS(GPSwhale21, ais_near_21, interpType = 'close', interpTime = 120) #empty table
gpsClose22 <- addAIS(GPSwhale22, ais_near_22, interpType = 'close', interpTime = 120)
gpsClose23 <- addAIS(GPSwhale23, ais_near_23, interpType = 'close', interpTime = 120)
```

Create column 'shipsin10km' that displays number of ships associated with timestamp
```{r}
# Viewing only ships within 10 km of buoys.
close2ships07 <- gpsClose07 %>% filter(shipDist < 10e3)
close2ships08 <- gpsClose08 %>% filter(shipDist < 10e3)
close2ships10 <- gpsClose10 %>% filter(shipDist < 10e3)
close2ships12 <- gpsClose12 %>% filter(shipDist < 10e3)
close2ships13 <- gpsClose13 %>% filter(shipDist < 10e3)
close2ships14 <- gpsClose14 %>% filter(shipDist < 10e3)
close2ships16 <- gpsClose16 %>% filter(shipDist < 10e3)
close2ships18 <- gpsClose18 %>% filter(shipDist < 10e3)
close2ships19 <- gpsClose19 %>% filter(shipDist < 10e3)
close2ships20 <- gpsClose20 %>% filter(shipDist < 10e3)
#close2ships21 <- gpsClose21 %>% filter(shipDist < 10e3) #not long enough of a buoy to detect
close2ships22 <- gpsClose22 %>% filter(shipDist < 10e3)
close2ships23 <- gpsClose23 %>% filter(shipDist < 10e3)

#Add column of ROUNDED UTC values to the lowest minute. Allows common field to bind
# Close2ships21 removed because no ships nearby
sslist <- c("close2ships07", "close2ships08", "close2ships10", "close2ships12", "close2ships13",
        "close2ships14", "close2ships16", "close2ships18", "close2ships19", "close2ships20",
        "close2ships22", "close2ships23")

for (sx in sslist) {
  dta <- get(sx)
  dta$UTCrnd <- lubridate::floor_date(as.POSIXct(dta$UTC), "minute")
  assign(sx, dta)
}

# Function that creates a column with number of ships within 10 km (shipin10km)
#Bind rows with ship nearby
#Summarize unique ships and display number of nearby vessels
ShpRows <- function(w, s) {
  z <- inner_join(w, s, by = join_by("UTC" == "UTCrnd")) %>%
    select(UTC, UTC.y, MMSI, vesselLength, vesselType, shipLat, shipLong, shipDist)%>%
    group_by(UTC) %>%
    summarize(shipin10km = n_distinct(MMSI))
}

```

Join to main df, add zeros to null values
```{r}
nearbyShp07 <- ShpRows(GPSwhale7, close2ships07)
GPSwhaleAIS7 <- left_join(GPSwhale7, nearbyShp07, by = join_by("UTC")) %>%
  mutate(shipin10km = coalesce(shipin10km, 0))

nearbyShp08 <- ShpRows(GPSwhale8, close2ships08)
GPSwhaleAIS8 <- left_join(GPSwhale8, nearbyShp08, by = join_by("UTC")) %>%
  mutate(shipin10km = coalesce(shipin10km, 0))

nearbyShp10 <- ShpRows(GPSwhale10, close2ships10)
GPSwhaleAIS10 <- left_join(GPSwhale10, nearbyShp10, by = join_by("UTC")) %>%
  mutate(shipin10km = coalesce(shipin10km, 0))

nearbyShp12 <- ShpRows(GPSwhale12, close2ships12)
GPSwhaleAIS12 <- left_join(GPSwhale12, nearbyShp12, by = join_by("UTC")) %>%
  mutate(shipin10km = coalesce(shipin10km, 0))

nearbyShp13 <- ShpRows(GPSwhale13, close2ships13)
GPSwhaleAIS13 <- left_join(GPSwhale13, nearbyShp13, by = join_by("UTC")) %>%
  mutate(shipin10km = coalesce(shipin10km, 0))

nearbyShp14 <- ShpRows(GPSwhale14, close2ships14)
GPSwhaleAIS14 <- left_join(GPSwhale14, nearbyShp14, by = join_by("UTC")) %>%
  mutate(shipin10km = coalesce(shipin10km, 0))

nearbyShp16 <- ShpRows(GPSwhale16, close2ships16)
GPSwhaleAIS16 <- left_join(GPSwhale16, nearbyShp16, by = join_by("UTC")) %>%
  mutate(shipin10km = coalesce(shipin10km, 0))

nearbyShp18 <- ShpRows(GPSwhale18, close2ships18)
GPSwhaleAIS18 <- left_join(GPSwhale18, nearbyShp18, by = join_by("UTC")) %>%
  mutate(shipin10km = coalesce(shipin10km, 0))

nearbyShp19 <- ShpRows(GPSwhale19, close2ships19)
GPSwhaleAIS19 <- left_join(GPSwhale19, nearbyShp19, by = join_by("UTC")) %>%
  mutate(shipin10km = coalesce(shipin10km, 0))

nearbyShp20 <- ShpRows(GPSwhale20, close2ships20)
GPSwhaleAIS20 <- left_join(GPSwhale20, nearbyShp20, by = join_by("UTC")) %>%
  mutate(shipin10km = coalesce(shipin10km, 0))

# Manually create column for buoys with no vessel activity
GPSwhaleAIS21 <- GPSwhale21 %>% add_column(shipin10km = 0)

nearbyShp22 <- ShpRows(GPSwhale22, close2ships22)
GPSwhaleAIS22 <- left_join(GPSwhale22, nearbyShp22, by = join_by("UTC")) %>%
  mutate(shipin10km = coalesce(shipin10km, 0))

nearbyShp23 <- ShpRows(GPSwhale23, close2ships23)
GPSwhaleAIS23 <- left_join(GPSwhale23, nearbyShp23, by = join_by("UTC")) %>%
  mutate(shipin10km = coalesce(shipin10km, 0))
```

Save new fulltrack rdas to D drive
```{r}
saveRDS(GPSwhaleAIS7, file = "data/GPSwhaleAIS7.rda")
saveRDS(GPSwhaleAIS8, file = "data/GPSwhaleAIS8.rda")
saveRDS(GPSwhaleAIS10, file = "data/GPSwhaleAIS10.rda")
saveRDS(GPSwhaleAIS12, file = "data/GPSwhaleAIS12.rda")
saveRDS(GPSwhaleAIS13, file = "data/GPSwhaleAIS13.rda")
saveRDS(GPSwhaleAIS14, file = "data/GPSwhaleAIS14.rda")
saveRDS(GPSwhaleAIS16, file = "data/GPSwhaleAIS16.rda")
saveRDS(GPSwhaleAIS18, file = "data/GPSwhaleAIS18.rda")
saveRDS(GPSwhaleAIS19, file = "data/GPSwhaleAIS19.rda")
saveRDS(GPSwhaleAIS20, file = "data/GPSwhaleAIS20.rda")
saveRDS(GPSwhaleAIS21, file = "data/GPSwhaleAIS21.rda")
saveRDS(GPSwhaleAIS22, file = "data/GPSwhaleAIS22.rda")
saveRDS(GPSwhaleAIS23, file = "data/GPSwhaleAIS23.rda")
```