### 2023-10-16 interpolation.Rmd
Create interpolation.Rmd
new rmd

---
title: "interpolation"
author: "Marina Bozinovic"
date: "2023-09-29"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
library(mgcv)
```

Read in rda
```{r}
# Read in original buoy 20 to compare
# Change name so as not to alter it
env20.orig <- readRDS(file = "D:/Buoy_dfs/GPSwhaleAIS20.rda")

# Interpolated buoy 20
env20.interp <- readRDS(file = "data/GPSwhaleAIS20_alt.rda")
```

Download and attach environmental variables to interpolated pts for buoy 20
```{r}
# Sea Surface Temp
env20.interp <- matchEnvData(env20.interp, nc = 'jplMURSST41', fileName = "D:/Environ_variables/analysedSST20_interp", var = c('analysed_sst')) # Download SST
env20.interp <- matchEnvData(env20.interp, nc = ss) %>%
  rename(., sst_mean = analysed_sst_mean) %>%
  dplyr::select(-matchLong_mean, -matchLat_mean, -matchTime_mean)  # Match SST

# Wind curl
env20.interp <- matchEnvData(env20.interp, nc = 'erdlasFnTran6_LonPM180', fileName = "D:/Environ_variables/windstress20_interp.nc", var = c('curl')) # Download wind curl
env20.interp <- matchEnvData(env20.interp, nc = w) %>%
     dplyr::select(-matchLong_mean, -matchLat_mean, -matchTime_mean)   # Match wind curl

# Distance to continental slope
bathy <- rast("D:/CCES_gebco_bathy.tif") # Bathymetry for all of CCE from https://download.gebco.net/
agg <- terra::aggregate(bathy, 5)  # Increase cell size by factor of 5 to be visible
cont <- st_as_sf(as.contour(bathy, )) # Create contour lines from bathymetry 
cont_slope <- cont %>% filter(level == -2000) # Establish continental shelf. Is 2000m the correct depth for continental shelf?
sf <- st_as_sf(env20.interp, coords = c("Longitude", "Latitude"), crs=4326, remove=F)
trk <- terra::rasterize(vect(sf), agg)
slop <- terra::distance(trk, vect(cont_slope))
ext <- terra::extract(slop, vect(sf))

dbind <- cbind(sf, ext) %>%
  dplyr::select(-ID) %>%
  rename("dist2slope" = last) %>%
  st_set_geometry(NULL) # remove geometry to prevent buoy dataframe from becoming sf object for future

# Depth
env20.interp <- matchEnvData(env20.interp, nc = 'srtm15plus', fileName = "D:/Environ_variables/depth20_alt.nc", var = c('z')) #Download depth
env20.interp <- matchEnvData(env20.interp, nc = dd) %>%
  rename(., depth = z_mean) %>%        # rename new column to "depth"
  mutate(depth = depth*-1) %>%
  dplyr::select(-matchLong_mean, -matchLat_mean, -matchTime_mean) #Match depth

# Bathymetric slope
bathy_slope <- terra::terrain(bathy, v = "slope")
slope <- terra::extract(bathy_slope, env20.interp[,3:2]) %>%  # Reads latitude and longitude of buoy dataframe
  dplyr::select(-ID) %>%
  rename("bathy_slope" = slope)
sbind <- cbind(env20.interp, slope)

# Distance to shore
env20.interp <- matchEnvData(env20.interp, nc = 'dist2coast_1deg', fileName = "D:/Environ_variables/dist2shore20_interp.nc", var = c('dist')) #Download bathy slope
env20.interp <- matchEnvData(env20.interp, nc = d2) %>%
  rename(., dist2shore = dist_mean) %>%
  dplyr::select(-matchLong_mean, -matchLat_mean, -matchTime_mean) # Match bathy slope

# Chlorophyll
env20.interp <- matchEnvData(env20.interp, nc = 'erdMH1chlamday', fileName = "D:/Environ_variables/chlorophyll20_interp.nc", var = c('chlorophyll')) #Download chl
env20.interp <- matchEnvData(env20.interp, nc = ch) %>%
  dplyr::select(-matchLong_mean, -matchLat_mean, -matchTime_mean) #Match chl

# Sea Surface Height
env20.interp <- matchEnvData(env20.interp, nc = 'nesdisSSH1day', fileName = "D:/Environ_variables/ssh20_interp.nc", var = c('sla')) #Download SSH
env20.interp <- matchEnvData(env20.interp, nc = sh) %>%
  rename(., ssh_mean = sla_mean) %>%
  dplyr::select(-matchLong_mean, -matchLat_mean, -matchTime_mean) # Match SSH

# Deep ocean variables
# Assumes themData function has run and exists in global environment
Htest <- matchEnvData(env20.interp, nc=PAMmisc::hycomList, var=c('water_temp', 'salinity'), raw=TRUE, depth = 0:1000)

thermData <- bind_rows(lapply(Htest, function(x) {
    # do thermo calcs
    result <- calcThermoVars(x$matchDepth, x$water_temp)
    # get specific values
    result$temp400 <- x$water_temp[x$matchDepth == 400]
    result$sal400 <- x$salinity[x$matchDepth == 400]
    result
}))

env20.interp <- cbind(env20.interp, thermData)

```

Save
```{r}
# To project
saveRDS(env20.interp, 'data/GPSwhaleAISenv20_interp.rda')

# To external hard drive
saveRDS(env20.interp, 'D:/Buoy_dfs/GPSwhaleAISenv20_interp.rda')
```

What is GAM of Buoy 20 without interpolation (original)
```{r}
trk20_noInterp <- gam(Wpresence>0 ~ depth + sst_mean + curl_mean + ttDepth + sal400
                      + TOL_125 + TOL_2000,
              data = env20.orig, family = binomial, method = "REML")

plot(trk20_noInterp, trans = plogis, all.terms = TRUE)
summary(trk20_noInterp)
# Env varb only: Deviance explained = 8.17%, R-sq = 0.051
# Env + SS variables: Deviance explained = 8.9%, R-sq = 0.0532
```

What is GAM of Buoy 20 with interpolation?
```{r}  
trk20_withInterp <- gam(Wpresence>0 ~ depth + sst_mean + curl_mean + ttDepth + sal400
                        + TOL_125 + TOL_2000,
              data = env20.interp, family = binomial, method = "REML")

plot(trk20_withInterp, trans = plogis, all.terms = TRUE)
summary(trk20_withInterp)
# Env varb only: Deviance explained = 7.91%, R-sq = 0.0634
# Env + SS variables: Deviance explained = 8.92%, R-sq = 0.0703
```

Summary Stats of Buoy 20 with/without interpolation
```{r}
# DF with removed questionable rows
#Before gap
#b4_mean <- sapply(env20.orig[830:849, 6:21], mean)
#b4_mean <- cbind(b4_mean, rownames(b4_mean))

#b4_sd <- sapply(env20.orig[830:849,6:21], sd)
#b4_sd <- cbind(b4_sd, rownames(b4_sd))

#b4_min <- sapply(env20.orig[830:849, 6:21], min)
#b4_min <- cbind(b4_min, rownames(b4_min))

#b4_max <- sapply(env20.orig[830:849, 6:21], max)
#b4_max <- cbind(b4_max, rownames(b4_max))
```

```{r}
#After gap
#aft_mean <- sapply(env20.orig[871:890, 6:21], mean)
#aft_mean <- cbind(aft_mean, rownames(aft_mean))

#aft_sd <- sapply(env20.orig[871:890, 6:21], sd)
#aft_sd <- cbind(aft_sd, rownames(aft_sd))

#aft_min <- sapply(env20.orig[871:890, 6:21], min)
#aft_min <- cbind(aft_min, rownames(aft_min))

#aft_max <- sapply(env20.orig[871:890, 6:21], max)
#aft_max <- cbind(aft_max, rownames(aft_max))
```

```{r}
b4af_mean <- sapply(env20.orig[840:860, 6:21], mean)
b4af_mean <- cbind(b4af_mean, rownames(b4af_mean))

b4af_sd <- sapply(env20.orig[840:860, 6:21], sd)
b4af_sd <- cbind(b4af_sd, rownames(b4af_sd))

b4af_min <- sapply(env20.orig[840:860, 6:21], min)
b4af_min <- cbind(b4af_min, rownames(b4af_min))

b4af_max <- sapply(env20.orig[840:860, 6:21], max)
b4af_max <- cbind(b4af_max, rownames(b4af_max))
```

```{r}
# DF of interpolated section
interp_mean <- sapply(env20.interp[850:870, 6:21], mean)
interp_mean <- cbind(interp_mean, rownames(interp_mean))

interp_sd <- sapply(env20.interp[850:870, 6:21], sd)
interp_sd <- cbind(interp_sd, rownames(interp_sd))

interp_min <- sapply(env20.interp[850:870, 6:21], min)
interp_min <- cbind(interp_min, rownames(interp_min))

interp_max <- sapply(env20.interp[850:870, 6:21], max)
interp_max <- cbind(interp_max, rownames(interp_max))
```

Create df for comparing before/after with interpolated points
```{r fig.width=10}
env20.orig <- env20.orig %>% mutate(dist2slopekm = dist2slope/1000,
                                  depthKm = depth/1000)

env20.interp <- env20.interp %>% mutate(dist2slopekm = dist2slope/1000,
                                  depthKm = depth/1000)

b4af_sample <- env20.orig[840:860, 6:24] %>%
  add_column(type = "beforeAfter")

interp_sample <- env20.interp[850:870, 6:24] %>%
  add_column(type = "interpolated")

compare_pts <- rbind(b4af_sample,interp_sample)

compare_longer1 <- pivot_longer(compare_pts, cols = c('sst_mean', 'mldTemp','mldTemp', 'ttTemp','temp400'),
             names_to = "variable",
             values_to = "values")

compare_longer2 <- pivot_longer(compare_pts, cols = c('TOL_125', 'TOL_2000', 'dist2slopekm'),
             names_to = "variable",
             values_to = "values")

compare_longer3 <- pivot_longer(compare_pts, cols = c('chlorophyll_mean', 'curl_mean', 'ssh_mean', 'depthKm'),
             names_to = "variable",
             values_to = "values")

compare_longer4 <- pivot_longer(compare_pts, cols = c('sal400', 'bathy_slope','ttDepth'),
             names_to = "variable",
             values_to = "values")

compare_longer5 <- pivot_longer(compare_pts, cols = c('mldDepth', 'ttDepth', 'depthKm'),
                                names_to = "variable",
                                values_to = "values")

depthKmplot <- pivot_longer(compare_pts, cols = 'depthKm',
                                names_to = "variable",
                                values_to = "values")

mldDepthplot <- pivot_longer(compare_pts, cols = 'mldDepth',
                                names_to = "variable",
                                values_to = "values")

ttDepthplot <- pivot_longer(compare_pts, cols = 'ttDepth',
                                names_to = "variable",
                                values_to = "values")
# Missing dist2shore because value does not change much
```

Plot
```{r fig.width=8}
ggplot(compare_longer1, aes(x = type, y = values, color = type)) + geom_boxplot() + 
  facet_wrap(~variable)
ggplot(compare_longer2, aes(x = type, y = values, color = type)) + geom_boxplot() + 
  facet_wrap(~variable)
ggplot(compare_longer3, aes(x = type, y = values, color = type)) + geom_boxplot() + 
  facet_wrap(~variable)
ggplot(compare_longer4, aes(x = type, y = values, color = type)) + geom_boxplot() + 
  facet_wrap(~variable)
#Depths only
ggplot(compare_longer5, aes(x = type, y = values, color = type)) + geom_boxplot() + 
  facet_wrap(~variable)

ggplot(depthKmplot, aes(x = type, y = values, color = type)) + geom_boxplot() +
  labs(title = "depth in Km")
ggplot(mldDepthplot, aes(x = type, y = values, color = type)) + geom_boxplot() +
  labs(title = "mldDepth")
ggplot(ttDepthplot, aes(x = type, y = values, color = type)) + geom_boxplot() +
  labs(title = "ttDepth")

```

Bind results & plot
```{r}
interp_results <- cbind(interp_mean, b4af_mean, interp_sd, b4af_sd, 
                        interp_min, b4af_min, interp_max, b4af_max) %>%
  format(scientific = F, digits = 1) %>%
  as.data.frame()

mean(as.numeric(interp_results$interp_sd))
mean(as.numeric(interp_results$b4af_sd))

#ggplot(interp_results) + geom_boxplot()

write.csv(interp_results, file = "interp_stats_drift20.csv")
```