### 2023-10-16 animations.Rmd
Create animations.Rmd

---
title: "animations"
author: "Marina Bozinovic"
date: "2023-02-17"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

### Map Animations

```{r}
# Install
#devtools::install_github("ropenscilabs/rnaturalearth")
#install.packages("rnaturalearth",
#                 repos = "http://packages.ropensci.org",
#                 type = "source")
#install.packages("gganimate")
#install.packages("gifski")
#install.packages("marmap")

# Libraries
library(sf)
library(tidyverse)
library(marmap)
library(gifski)
library(gganimate)
library(rnaturalearth)
```

### New ggplot wtih bplot as basemap

```{r}
library(mapproj)
drift <- track_BB_OL_10

# Get bathymetry (code from Ocean Tracking Network #https://www.youtube.com/watch?v=xT2F-Y9bKLk)
bgo <- getNOAA.bathy(lon1 = min(drift$long-0.7), lon2 = max(drift$long + 0.7),
                     lat1 = min(drift$lat-0.7), lat2=max(drift$lat + 0.7), resolution = 1)

# Create basemap
#bf <- fortify.bathy(bgo, geom=c("r", "c")) + scale_fill_etopo() # converts bathy object to data frame 
# ABOVE DOESNT WORK
bplot <- autoplot(bgo, geom = c("raster", "contour"), coast = TRUE) + 
  scale_fill_etopo() #creates ggplot object

# Plot drift
p <- bplot + 
  geom_path(data = drift, 
            aes(x=long,y=lat,group=spotID), 
            alpha = 0.5) +
  geom_point(data = drift, 
             aes(x=long,y=lat,group=spotID, size = OL_125, color = OL_125),
             shape=19, stroke = 1) +

  #formatting  
  scale_color_viridis_c(option = "inferno") +
  scale_size_continuous(range = c(0.1,10)) +
  labs(x=NULL, y=NULL, 
       fill = 'Depth',
       color = 'OL 125')+
  theme(panel.grid = element_blank())

# Plot AIS
b <- bplot + 
  geom_point(data = AIS_sf, mapping = aes(x=LON,y=LAT, group = VesselName)) +
  scale_x_continuous(limits = c(min(drift$long-0.7), max(drift$long + 0.7))) + 
  scale_y_continuous(limits = c(min(drift$lat-0.7), max(drift$lat + 0.7))) +
  theme(panel.grid = element_blank())

# Plot AIS + drift
z <- bplot +
  geom_point(data = AIS_sf, mapping = aes(x=LON,y=LAT, group = VesselName)) +
  geom_path(data = drift, 
            aes(x=long,y=lat,group=spotID), 
            alpha = 0.5) +
  geom_point(data = drift, 
             aes(x=long,y=lat,group=spotID, size = OL_125, color = OL_125),
             shape=19, stroke = 1) +
  #formatting  
  scale_color_viridis_c(option = "inferno") +
  scale_size_continuous(range = c(0.1,10)) +
  labs(x=NULL, y=NULL, 
       fill = 'Depth',
       color = 'OL 125')+
  scale_x_continuous(limits = c(min(drift$long-0.7), max(drift$long + 0.7))) + 
  scale_y_continuous(limits = c(min(drift$lat-0.7), max(drift$lat + 0.7)))
```

```{r}
# Animate drift
anim <- p +
  transition_reveal(along = dateTime)+
  ease_aes('linear')+
  shadow_wake(0.1) +
  ggtitle("Date: {frame_along}") +

# Animate with control over frames per second and number of frames
animate(anim, nframes = 70, fps = 3)
anim_save("gifs/drift10.gif")

# Animate AIS. Need to continue working on this.
anim <-  b +
  transition_reveal(BaseDateTime) +
  ease_aes('linear') +
  shadow_wake(wake_length = 0.5) +
  ggtitle("Date: {frame_along}")

# Animate with control over frames per second and number of frames
animate(anim, nframes = 400, fps = 3)

```

### Plotting whale data

```{r}
# shorten time range
#whales10 <- whales10 %>% filter(StartTime > "2018-10-05 00:00:00")

# too smooth, doesnt work.
d <- bplot +
  geom_point(data = whales10, aes(x = as.numeric(Longitude), y = as.numeric(Latitude)), color = "cyan") +
  ggtitle("Drift 10: Whale detections")

anim <-  d +
  transition_reveal(along = StartTime) +
  exit_shrink() +
  ggtitle("Date: {frame_along}")

animate(anim, nframes = 400, fps = 3)

## The issue with showing drift path and whale detections is that gganimate only allows one type of transition for the whole plot. May need to join drift and whale detections and include null values?
```

### Join whale detections with drift 10

```{r}
w_d <- left_join(drift, whales10, by = "lat", keep = TRUE)

w_d_plt <- ggplot() +
  geom_path(data = w_d, aes(x=lon, y=lat.x)) +
  geom_point(data = w_d, aes(x=lon, y = lat.x, color = "species"))

w_d_plt + 
  transition_reveal(along = dateTime) +
  ease_aes() +
  shadow_trail(distance = 0.7) +
  ggtitle("Date:{frame_along}")

## Needs shorter time frame

```